### 2024-03-19 notes_fricatives.Rmd
update notes_fricatives + papers

# Notes for the introduction
Catalan and Spanish are two languages that have coexisted for centuries. 
<https://books.google.com/books?id=AhL9WsTaNzYC&printsec=frontcover&hl=es&vq=%22Estudis+de+fon%C3%A8tica+experimental+del+catal%C3%A0+oriental+central%22&source=gbs_citations_module_r&cad=1#v=onepage&q=%22Estudis%20de%20fon%C3%A8tica%20experimental%20del%20catal%C3%A0%20oriental%20central%22&f=false>

1. Catalan coexists with Spanish 
2. There is a bidirectional influence between these two languages
3. Catalan influences Spanish and Spanish influences Catalan in all linguistic levels (possible citation: Davidson, 2014, 2015), although this influence is probably more notable in the production of sounds (citations?).

# Notes on alveolar fricatives
Both Catalan and Peninsular Spanish sound inventories have an phoneme **[s]/s/**, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015) **(CITATIONS)**. In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p. 279). In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [sɛl] - *zel* [zɛl] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019). Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022). For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tórəs] and *torres antigues* [torəzəntiɣəs].

Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these **phonemes/allophones** merged through a devoicing process (Hualde & Colina, 2015). Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020). Hence, Spanish inventory does not have the phoneme **[z]/z/**, but **[z]/z/** may occur as an allophone of **[s]/s/**, due to a voice assimilation process. It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules (Institut d'Estudis Catalans, 2022), the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas (Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268). The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia (Davidson, 2020).

In summary, both Catalan and Spanish have a voiceless alveolar fricative in their inventories. Catalan also has a voiced variant, which is produced when this sound is followed by a voiced consonant or a vowel. Spanish might produced this sound only when it is followed by a voiced consonant, although /s/ might be "partially voiced" (Hualde & Colina, 2015). Thus, these phonemes, although they have the same allophones, their contextual distribution is different (Llisterri, 2023). Moreover, some studies have shown that there is a bidirectional influence between both languages in the production of the voiced vs the voiceless alveolar fricative, and that such influence might be different based on the context. These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.

4. Both Spanish and Catalan sound inventories have an apical-alveolar voiceless /s/, "which is articulated with a gesture of the tongue-tip toward the alveolar ridge" (Davidson, 2014)
5. Most of the Catalan varieties have "an additional apical-alveolar phoneme, voiced /z/" 
6. Linguistic constraints of /z/ -> ? 
7. Minimal pairs such as *casa* (/z/) and *caça* (/s/) -> Este mismo ejemplo sale en Davidson (2014). ¿Buscar otro?
8. /z/ - /s/ contrast is neutralized at the end of the word. The production of the voiced or voiceless apical-alveolar depends on the following sound, in the following word. For example, when it is followed by a vowel, the alveolar fricative in coda position (does it make sense to say coda position here?) is voiced: *torres* ([s]) and *torres antigues* ([z])
9. Intervocalic [z] is a common and systematic feature of Catalan, but it has not been attested as a systematic feature in monolingual varieties of Spanish (Davidson, 2014, 2015) -> Hence, un problemaaaaa :) 
10. Davidson (2014, p. 6): "We may expect that CCS voicing be more strongly disfavored word-medially because this position is a site of phonemic voicing contrast in Catalan, whereas the word-final position, being a site of purely phonetic voicing for both languages (albeit systematic in Catalan), may constitute a better candidate for crosslinguistic phonetic transfer from Catalan to CCS."

Cited articles:
Benet, A.; Cortés, S., & Lleó, C. (2012). Devoicing of sibilants as a segmental cue to the influence of Spanish onto current Catalan phonology. *Hamburg Studies on Multilingualism*. (book chapter)

Citació: Carrera-Sabaté, Josefina  & Prieto, Pilar (coords.) (2012-2019)  
Guies de Pronunciació del català. Pàgina web: <http://www.ub.edu/guiesdepronunciacio>.

Davidson, J. (2020).

Hualde & Colina. (2015). Los sonidos del español

Institut d’Estudis Catalans (2022). Gramàtica bàsica i d’ús de la llengua catalana. <https://gbu.iec.cat/inici>

Llisterri, J. (2023). La fonética contrastiva. <https://joaquimllisterri.cat/applied_linguistics/L2_phonetics/Fon_Cont.html>

Lloret, MR. (2011). La fonologia del català. Santillana.

Recasens Vives, D. (1996). *Fonètica descriptiva del català: Assaig de caracterització de la pronúncia del vocalisme i el consonantisme del català al segle XX*. Institut d'Estudis Catalans. (book) <https://books.google.com/books?id=2UCDqrfRfVcC&printsec=frontcover&source=gbs_ge_summary_r&cad=0#v=onepage&q&f=false>

Other useful books/articles:
Prieto 2004, 204-205 -> Prieto, Pilar. 2004. Fonètica i fonologia: Els sons del català [Phonetics and Phonology: The Sounds of Catalan]. Barcelona: Editorial UOC.

Quilis 1981, 234-235 -> Quilis, Antonio. 1981. Fonética acústica de la lengua española [Acoustic Phonetics of the Spanish Language]. Madrid: Gredos.

Recasens 1991, 267-268 -> Recasens, Daniel. 1986. Estudis de fonètica experimental del català oriental central [Experimental Phonetic Studies on East-Central Catalan]. Barcelona: Publicacions de L’Abadia de Montserrat. ---------. 1991. Fonètica descriptiva del català [Descriptive Phonetics of Catalan]. Barcelona: Institut d’Estudis Catalans. ---------. 1993. Fonètica i fonologia [Phonetics and Phonology]. Barcelona: Enciclopèdia Catalana.

Martínez Celdrán & Fernández Planas 2007 -> Martínez Celdrán, Eugenio, and Ana María Fernández Planas. 2007. Manual de fonética española: Articulaciones y sonidos del español [Manual of Spanish Phonetics: Articulations and Sounds of Spanish]. Barcelona: Ariel.

Julià i Muné 2008, 66-67 -> Julià i Muné, Joan. 2008. “Els sons del català [The Sounds of Catalan].” In Gramàtica del Català Contemporani, Volum 1: Introducció, Fonètica i Fonologia, Morfologia, ed. by Joan Solà, Maria-Rosa Lloret, Joan Mascaró, and Manuel Pérez Saldanya, 37-87. Barcelona: Editorial Empúries.

Arnal’s (2011, p. 15) -> Arnal, A. (2011). Linguistic Changes in the Catalan Spoken in Catalonia under New Contact Conditions. Journal of Language Contact, 4, 5-25

Woolard’s 2007 ->

Lleó 2006 -> Lleó, C. 2006. Fenòmens evolutius i d’ús oral del català des de la finestra psicolingüística de l’adquisició bilingüe. In Actes del 13è Col·loqui internacional de llengua i literatura catalanes (AILLC), eds. S. Martí, M. Cabré, F. Feliu, N. Iglesias & D. Prats, 361–385. Barcelona: Publicacions de l’Abadia de Montserrat.

### 2024-03-20 notes_fricatives.Rmd
manuscript

@@
 # Notes on alveolar fricatives
-Both Catalan and Peninsular Spanish sound inventories have an phoneme **[s]/s/**, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015) **(CITATIONS)**.
+Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015).
 In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p.
@@
 
-Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these **phonemes/allophones** merged through a devoicing process (Hualde & Colina, 2015).
+Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process (Hualde & Colina, 2015).
 Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020).
-Hence, Spanish inventory does not have the phoneme **[z]/z/**, but **[z]/z/** may occur as an allophone of **[s]/s/**, due to a voice assimilation process.
+Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process.
 It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).

### 2024-03-20 manuscript/main.Rmd

---
documentclass: article
classoption: [a4paper,11pt,twocolumn]
subparagraph: yes
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1]{fontenc}
  #- \usepackage[leqno,fleqn]{amsmath}
output:
  pdf_document:
    highlight: zenburn
    includes:
      before_body: includes/tex/doc_prefix.tex
      in_header: includes/tex/header.tex
    keep_tex: yes
    number_sections: yes
bibliography: includes/bib/icphs2023.bib
---

```{r, 'setup', include=FALSE}
library(knitr)
f <- function(x, options) {
  lab <- paste0(options$fig.lp, options$label)
  paste(
        "\\begin{figure}[!h]\n\\caption{", options$fig.cap.top, "}\\label{", lab,"}\n", 
        hook_plot_tex(x, options), 
        "\\end{figure}\n\n", sep = "")
}
knit_hooks$set(plot = f)
```

\section{Introduction}

Write something here

\begin{itemize}
  \item they must be written in English
  \item they must not contain the name(s) of the author(s) (for
anonymous review)
  \item the maximum is four pages for Congress papers, plus up to one additional page of references. Papers for Plenary lectures can be longer
  \item they must be submitted in PDF format (cf.\ Section~4)
  \item the paper submission will occur via a web interface
\end{itemize}
This paper template can be found on the conference website. If there
are special questions or wishes regarding paper preparation and
submission for ICPhS 2023, correspondence should be addressed to
icphs2023@guarant.cz. Information for full paper submission will be
available on the web at \url{https://www.icphs2023.org/}.

\section{Background Literature}

Write something?

\subsection{The production of the alveolar fricative in Catalan and Spanish}

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015). In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p. 279). In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [sɛl] - *zel* [zɛl] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019). Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022). For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tórəs] and *torres antigues* [torəzəntiɣəs].

Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process (Hualde & Colina, 2015). Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020). Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process. It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules (Institut d'Estudis Catalans, 2022), the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas (Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268). The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia (Davidson, 2020).

In summary, both Catalan and Spanish have a voiceless alveolar fricative in their inventories. Catalan also has a voiced variant, which is produced when this sound is followed by a voiced consonant or a vowel. Spanish might produced this sound only when it is followed by a voiced consonant, although /s/ might be "partially voiced" (Hualde & Colina, 2015). Thus, these phonemes, although they have the same allophones, their contextual distribution is different (Llisterri, 2023). Moreover, some studies have shown that there is a bidirectional influence between both languages in the production of the voiced vs the voiceless alveolar fricative, and that such influence might be different based on the context. These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.

\subsubsection{Headings}

Section headings are centered in boldface with capitalized letters. Sub-headings start at the left margin in the column with the first letter capitalized and the rest of the heading in lower case. Sub-sub-headings appear like sub-headings, except that they are in italics and not boldface. See examples in this file. No more than 3 levels of headings should be used.  Empty lines should be left above and below each section heading.

\subsection{Text font}

Times or Times New Roman font is used for the main text. Recommended font size is 11 points. Other font types may be used if needed for special purposes. If using any non-Unicode fonts, these must be embedded in the final PDF file.

The \LaTeX{} template can be used either with plain \LaTeX{} or \XeLaTeX.

\subsection{Figures}

All figures should be centered on the column (or page, if the figure
spans both columns). Figure captions should follow each figure and
have the format given in Fig. \ref{fig:vowels}.

Figures should preferably be line drawings. If they contain grey
shades or colours, it should be checked that they print well on a
high-quality noncolour laser printer.

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=6cm]{./includes/figures/ipa.eps}
\caption{The vowel chart used in the International Phonetic
Alphabet (IPA).}\label{fig:vowels}
\end{center}
\end{figure}

\subsection{Tables}

An example of a table is shown as Table~\ref{tab:decibel}.  Somewhat
different styles are allowed according to the type and purpose of the
table. Colour should not be used, but grey shading is allowed. There
should be a margin of 6~points (pt) above and below the table.

The caption text may be above or below the table, but this should be
consistent throughout the submission. Left and right indentation of
the caption should be 0.5~cm.

\begin{table}[!ht]
  \begin{center}
  \begin{tabular}{|c|c|}
  \hline
  \rowcolor[gray]{.75}
  ratio    & Decibels\\
  \hline
  $1/1$    & $0$\\
  $2/1$    & $6$\\
  $3.16$   & $10$\\
  $1/10$   & $-20$\\
  $10/1$   & $20$\\
  $100/1$  & $40$\\
  $1000/1$ & $60$\\
  \hline
  \end{tabular}
  \caption{This is an example of a table showing Decibel (dB)
  ratios.}\label{tab:decibel}
  \end{center}
\end{table}

\subsection{Equations}

Equations should be placed on separate lines and numbered. An example
of an equation is
\begin{equation}\label{eq:tzero}
T_0 = \frac{1}{f_0}.
\end{equation}
Numbers of equations can be on the right or on the left margin of the
text column.

\subsection{Examples}

Examples from other languages can either be presented in the body text, or, if referred to elsewhere or particularly long and complex, can be put on a separate, numbered line, as should be done for equations.

\subsection{Phonetic fonts}

We recommend that you use the \textsc{tipa} package for IPA phonetic symbols. For information about how to access Unicode fonts in the Word template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}. The font you use must be embedded. Please remember to check this,  e.g. by inspecting the ``Document Properties --- Fonts'' in Acrobat Reader.

\subsection{Page numbering}

Page numbers will be added electronically to the document
later. \textit{Please do not add page numbers and please do not insert
any footers or headers!}

\subsection{References}

Please use just the reference number in square brackets. Formulations
with author names like "\ldots\ as Ladefoged \cite{Ladefoged:2003}
showed that \ldots" are acceptable but not "as shown in [Ladefoged,
3]" or "as shown in (Ladefoged [3])".

References are to be numbered in alphabetical order. Please
double-check the final version of your paper with regard to the
correct correspondence of references to their numbers.

\subsection{Hyperlinks}

Links to URLs or email addresses should be formatted as normal text,
\textit{not} as hyperlinks and not blue or underlined etc. Usually
hyperlinks to web pages are listed in the references section. If
required, line breaks can be placed within URLs after slashes or
dashes (cf.~\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that no hyphens
are inserted.

\subsection{Footnotes and endnotes}

If footnotes cannot be avoided they should appear as
endnotes.\footnote{This footnote appears here as an endnote.}

\section{Multimedia files}

Multimedia data that are part of the paper are to be embedded in the submitted PDF; they cannot be submitted as supplementary data. Any images are to be included in the paper as Figures (see Section 2.3 above). It is the authors' responsibility to check image quality ahead of submission. Audio examples are to be embedded within the PDF. To do this, authors can generate the PDF, and then embed the audio files using Adobe Acrobat Professional, or other software that offers the same outcome, so that the audio is included in the PDF. The presence of audio data should be identified in the text.

We encourage authors to illustrate video data using still photographs from the video, and to include them as figures in the PDF.  We cannot accept embedded video files, but authors are welcome to refer readers to a URL on the internet where these can be viewed.

\section{PDF details}

PDF files submitted must comply with the following requirements:

\begin{enumerate}
\item all special fonts and symbols must be embedded in the PDF
  file so that correct rendering of the PDF does not depend on the
  fonts installed on the viewer's computer
\item there must be no
  password protection on the PDF file, i.e. PDF files must not be
  protected by PDF security in any way, i.e. content extraction,
  document assembly, high resolution printing etc. must not be
  forbidden
\item PDF files should not contain any colours, hyperlinks,
  multimedia or 3D content, and no JavaScript or forms
\item PDF files
  should be no larger than 5 MB.
\end{enumerate}

\section{Anonymity}

In ICPhS 2023 submissions, an anonymous reviewing process will be used. This means that for the first submission the name(s) of the author(s) and
their affiliation(s) \emph{must not} be mentioned. In addition, please refrain from using acknowledgements.  Please also try to make your own previous research as anonymous as possible. As an example: do not write "In our previous study \cite{Stevens:1999} we could show\ldots" but "As shown in \cite{Stevens:1999}\ldots". Or refer to your own published or otherwise widely known work, and to that of the other authors, in the "Julius Caesar style", i.e. in the third person (for example: his work, her work, their work). Reference as "anonymous" only work that you or the other authors have submitted for publication, but which has not yet been published, e.g. \cite{Anonymous:submitted}.

Please make sure that no author details appear in the Document Properties of the PDF file. \emph{For the revised paper submission author details are of course needed}. Acknowledgements and references to one's own work are possible as usual.

\section{Format of references}

References are formatted using the IEEE standard (available in various reference management systems like Zotero or Mendeley). If you do not use a reference management system, please use the examples provided for monographs \cite{Fant:1960}, contributions to volumes \cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles in conference proceedings \cite{Ladefoged:2003}. Abbreviations of
well-known conferences and journals are possible \cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016} are referenced according to authors' instructions.

\section{R scripts}

You can use knitr code chunks like in any .Rmd document:

```{r, 'chunk_ex'}
2 + 2
```

This includes inline code: 2 + 2 = `r 2 + 2`.

R scripts can be loaded as well:

```{r, 'load_script', echo=TRUE, message=FALSE, warning=FALSE}
source("./includes/scripts/analysis.R")
```

Which means you have access to any objects assigned in the script, like Figure 
\ref{fig:ex_plot}:

```{r, 'ex_plot', fig.align='center', fig.width=2.75, fig.height=1.5, fig.cap.top="This is a figure caption.", fig.pos='h', fig.lp='fig:'} 
cars_plot
```

\bibliographystyle{./includes/bib/IEEEtran.bst}
\bibliography{./includes/bib/icphs2023.bib}

\theendnotes

### 2024-03-20 manuscript/main.tex

--- 2024-03-20 manuscript/main.Rmd
@@
----
-documentclass: article
-classoption: [a4paper,11pt,twocolumn]
-subparagraph: yes
-header-includes:
-  - \usepackage[utf8]{inputenc}
-  - \usepackage[T1]{fontenc}
-  #- \usepackage[leqno,fleqn]{amsmath}
-output:
-  pdf_document:
-    highlight: zenburn
-    includes:
-      before_body: includes/tex/doc_prefix.tex
-      in_header: includes/tex/header.tex
-    keep_tex: yes
-    number_sections: yes
-bibliography: includes/bib/icphs2023.bib
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  a4paper,
+  11pt,
+  twocolumn]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{48,48,48}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
+\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{\textbf{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.87,0.87,0.75}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.86,0.86,0.80}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.76,0.75,0.62}{#1}}
+\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.75,0.75,0.82}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
+\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
+\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.94,0.94,0.82}{#1}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{\textbf{#1}}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+%------------------------------------------------------------------------------%
+% PAPER TEMPLATE FOR ICPHS 2023 Prague                                         %
+%                                                                              %
+% Original template downloaded from:                                           %
+% http://www.icphs2023.org/call-for-papers/                                    %
+%                                                                              %
+% Reformatted to work with Rmarkdown and R by:                                 %
+% Joseph V. Casillas | Rutgers Univesity |11/11/2022                           %
+%                                                                              %
+% Available for download at:                                                   %
+% https://github.com/jvcasillas/icphs2023_rmd_template                         %
+%------------------------------------------------------------------------------%
 
-```{r, 'setup', include=FALSE}
-library(knitr)
-f <- function(x, options) {
-  lab <- paste0(options$fig.lp, options$label)
-  paste(
-        "\\begin{figure}[!h]\n\\caption{", options$fig.cap.top, "}\\label{", lab,"}\n", 
-        hook_plot_tex(x, options), 
-        "\\end{figure}\n\n", sep = "")
-}
-knit_hooks$set(plot = f)
-```
+% Packages
+\usepackage{./includes/tex/icphs2023}
+\usepackage{metalogo} 
+\usepackage{epstopdf}
+\usepackage{tipa}
+
+% Links and urls must be black
+%\hypersetup{urlcolor=black, citecolor=black, linkcolor=black}
+
+% Packages removed from icphs2023.sty because of conflicts
+% They have been added to the .Rmd yaml front matter
+% \usepackage[latin1]{inputenc}
+% \usepackage[T1]{fontenc}
+% \usepackage[leqno,fleqn]{amsmath}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\author{}
+\date{\vspace{-2.5em}}
+
+\begin{document}
+
+\title{Paper template for {ICPhS} 2023 Prague}
+\author{Please write XXX instead of the name(s) of the author(s)}
+\organization{Please write XXX instead of the affiliation(s)}
+\email{please write XXX instead of the email address(es)}
+
+\maketitle
+
+\begin{abstract}
+This is the layout specification and template definition for the papers of ICPhS XX (the 20th International Congress of Phonetic Sciences, which will be held in Prague, Czech Republic, August 7-11, 2023).
+This template is a revised version of the one used for Melbourne ICPhS XIX in 2019, originally generated from the template for Speech Prosody 2006 in Dresden.
+
+The abstract may consist of more than one paragraph but must be kept within a 150-word limit. 
+This abstract will be printed in the abstract booklet to be given out at the conference.
+\end{abstract}
+
+\keywords{There is space for up to five self-selected keywords (maximally two 
+lines).}
 
@@
 \end{itemize}
-This paper template can be found on the conference website. If there
-are special questions or wishes regarding paper preparation and
-submission for ICPhS 2023, correspondence should be addressed to
-icphs2023@guarant.cz. Information for full paper submission will be
-available on the web at \url{https://www.icphs2023.org/}.
+
+This paper template can be found on the conference website. If there are
+special questions or wishes regarding paper preparation and submission
+for ICPhS 2023, correspondence should be addressed to
+\href{mailto:icphs2023@guarant.cz}{\nolinkurl{icphs2023@guarant.cz}}.
+Information for full paper submission will be available on the web at
+\url{https://www.icphs2023.org/}.
 
@@
 
-Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015).
-In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p.
-279).
-In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [sɛl] - *zel* [zɛl] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019).
-Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022).
-For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tórəs] and *torres antigues* [torəzəntiɣəs].
+Both Catalan and Peninsular Spanish sound inventories have an phoneme
+/s/, which can be classified as a voiceless alveolar fricative (Hualde
+\& Colina, 2015). In addition, most of the Catalan varieties have a
+voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens
+Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
+phonemically, which leads to minimal pairs like \emph{cel} {[}sɛl{]} -
+\emph{zel} {[}zɛl{]} and \emph{fossa} {[}fósa{]} - \emph{fosa}
+{[}fóza{]} (Carrera-Sabaté \& Prieto, 2012-2019). Conversely, at the end
+of the word, this contrast is neutralized, being the voicing or
+devoicing of the alveolar fricative dependent on the following sound
+(Institut d'Estudis Catalans, 2022). For example, if it is followed by a
+vowel, the alveolar fricative is voiced: \emph{torres} {[}tórəs{]} and
+\emph{torres antigues} {[}torəzəntiɣəs{]}.
 
-Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process (Hualde & Colina, 2015).
-Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020).
-Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process.
-It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).
+Interestingly, Medieval Spanish, similarly to Modern Catalan,
+distinguished between /s/ and /z/ in intervocalic position within a
+word, although these sounds merged through a devoicing process (Hualde
+\& Colina, 2015). Nevertheless, in those modern varieties where the /s/
+in coda position is produced, such as in Peninsular Spanish, this
+voiceless phoneme might assimilate to the following sound, especially if
+it is a voiced consonant (\emph{mismo}, {[}mízmo{]}), rather than a
+vowel (\emph{las alas}, {[}lasálas{]}) (Hualde \& Colina, 2015; Hualde
+et al., 2020). Hence, Spanish inventory does not have the phoneme /z/,
+but {[}z{]} may occur as an allophone of /s/, due to a voice
+assimilation process. It is important to mention, though, that it has
+been claimed that this voicing is generally not complete, but partial
+(Benet et al., 2012; Hualde \& Colina, 2015; Hualde et al., 2020).
 
-There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules (Institut d'Estudis Catalans, 2022), the [z] should be produced, but [s] is produced instead.
-This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas (Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
-268).
-The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia.
-Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia (Davidson, 2020).
+There is some evidence of the influence of Spanish on the production of
+the Catalan alveolar fricative, in contexts where, according to
+prescriptivist rules (Institut d'Estudis Catalans, 2022), the {[}z{]}
+should be produced, but {[}s{]} is produced instead. This phenomenon has
+mainly been attested in those areas where there has been an extended
+contact between Catalan and Spanish, such asthe so-called ``franja de
+Aragón'' (borderland between Catalonia and Aragon), some areas of the
+Valencian Community, and Barcelona and other urban areas (Benet et al.,
+2012; Davidson, 2020; Recasens Vives, 1996, p.~268). The opposite
+pattern has also been observed - that is, the voicing of the alveolar
+fricative in the Spanish spoken in Barcelona and Valencia.
+Interestingly, although in both Barcelona and Valencia Catalan and
+Spanish coexist and influence each other, it seems that Catalan has a
+greater influence on Spanish in Barcelona, but Spanish has a greater
+influence on Catalan in Valencia (Davidson, 2020).
 
-In summary, both Catalan and Spanish have a voiceless alveolar fricative in their inventories.
-Catalan also has a voiced variant, which is produced when this sound is followed by a voiced consonant or a vowel.
-Spanish might produced this sound only when it is followed by a voiced consonant, although /s/ might be "partially voiced" (Hualde & Colina, 2015).
-Thus, these phonemes, although they have the same allophones, their contextual distribution is different (Llisterri, 2023).
-Moreover, some studies have shown that there is a bidirectional influence between both languages in the production of the voiced vs the voiceless alveolar fricative, and that such influence might be different based on the context.
-These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.
+In summary, both Catalan and Spanish have a voiceless alveolar fricative
+in their inventories. Catalan also has a voiced variant, which is
+produced when this sound is followed by a voiced consonant or a vowel.
+Spanish might produced this sound only when it is followed by a voiced
+consonant, although /s/ might be ``partially voiced'' (Hualde \& Colina,
+2015). Thus, these phonemes, although they have the same allophones,
+their contextual distribution is different (Llisterri, 2023). Moreover,
+some studies have shown that there is a bidirectional influence between
+both languages in the production of the voiced vs the voiceless alveolar
+fricative, and that such influence might be different based on the
+context. These divergent patterns might be explained by the different
+sociopolitic and linguistic realities of each city, and might have an
+effect on speakers' attitudes towards the devoicing of the alveolar
+fricative /z/ in Catalan.
 
@@
 Section headings are centered in boldface with capitalized letters.
-Sub-headings start at the left margin in the column with the first letter capitalized and the rest of the heading in lower case.
-Sub-sub-headings appear like sub-headings, except that they are in italics and not boldface.
-See examples in this file.
-No more than 3 levels of headings should be used.
-Empty lines should be left above and below each section heading.
+Sub-headings start at the left margin in the column with the first
+letter capitalized and the rest of the heading in lower case.
+Sub-sub-headings appear like sub-headings, except that they are in
+italics and not boldface. See examples in this file. No more than 3
+levels of headings should be used. Empty lines should be left above and
+below each section heading.
 
@@
 
-Times or Times New Roman font is used for the main text.
-Recommended font size is 11 points.
-Other font types may be used if needed for special purposes.
-If using any non-Unicode fonts, these must be embedded in the final PDF file.
+Times or Times New Roman font is used for the main text. Recommended
+font size is 11 points. Other font types may be used if needed for
+special purposes. If using any non-Unicode fonts, these must be embedded
+in the final PDF file.
 
-The \LaTeX{} template can be used either with plain \LaTeX{} or \XeLaTeX.
+The \LaTeX{} template can be used either with plain \LaTeX{} or
+\XeLaTeX.
 
@@
 All figures should be centered on the column (or page, if the figure
-spans both columns). Figure captions should follow each figure and
-have the format given in Fig. \ref{fig:vowels}.
+spans both columns). Figure captions should follow each figure and have
+the format given in Fig. \ref{fig:vowels}.
 
-Figures should preferably be line drawings. If they contain grey
-shades or colours, it should be checked that they print well on a
-high-quality noncolour laser printer.
+Figures should preferably be line drawings. If they contain grey shades
+or colours, it should be checked that they print well on a high-quality
+noncolour laser printer.
 
@@
 
-An example of a table is shown as Table~\ref{tab:decibel}.  Somewhat
-different styles are allowed according to the type and purpose of the
-table. Colour should not be used, but grey shading is allowed. There
-should be a margin of 6~points (pt) above and below the table.
+An example of a table is shown as
+Table\textasciitilde{}\ref{tab:decibel}. Somewhat different styles are
+allowed according to the type and purpose of the table. Colour should
+not be used, but grey shading is allowed. There should be a margin of
+6\textasciitilde points (pt) above and below the table.
 
 The caption text may be above or below the table, but this should be
-consistent throughout the submission. Left and right indentation of
-the caption should be 0.5~cm.
+consistent throughout the submission. Left and right indentation of the
+caption should be 0.5\textasciitilde cm.
 
@@
 
-Equations should be placed on separate lines and numbered. An example
-of an equation is
-\begin{equation}\label{eq:tzero}
+Equations should be placed on separate lines and numbered. An example of
+an equation is \begin{equation}\label{eq:tzero}
 T_0 = \frac{1}{f_0}.
-\end{equation}
-Numbers of equations can be on the right or on the left margin of the
-text column.
+\end{equation} Numbers of equations can be on the right or on the left
+margin of the text column.
 
@@
 
-Examples from other languages can either be presented in the body text, or, if referred to elsewhere or particularly long and complex, can be put on a separate, numbered line, as should be done for equations.
+Examples from other languages can either be presented in the body text,
+or, if referred to elsewhere or particularly long and complex, can be
+put on a separate, numbered line, as should be done for equations.
 
@@
 
-We recommend that you use the \textsc{tipa} package for IPA phonetic symbols.
-For information about how to access Unicode fonts in the Word template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}.
-The font you use must be embedded.
-Please remember to check this,  e.g.
-by inspecting the ``Document Properties --- Fonts'' in Acrobat Reader.
+We recommend that you use the \textsc{tipa} package for IPA phonetic
+symbols. For information about how to access Unicode fonts in the Word
+template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}. The font you use
+must be embedded. Please remember to check this, e.g.~by inspecting the
+``Document Properties --- Fonts'\,' in Acrobat Reader.
 
@@
 
-Page numbers will be added electronically to the document
-later. \textit{Please do not add page numbers and please do not insert
+Page numbers will be added electronically to the document later.
+\textit{Please do not add page numbers and please do not insert
 any footers or headers!}
@@
 Please use just the reference number in square brackets. Formulations
-with author names like "\ldots\ as Ladefoged \cite{Ladefoged:2003}
-showed that \ldots" are acceptable but not "as shown in [Ladefoged,
-3]" or "as shown in (Ladefoged [3])".
+with author names like ``\ldots~as Ladefoged \cite{Ladefoged:2003}
+showed that \ldots'' are acceptable but not ``as shown in {[}Ladefoged,
+3{]}'' or ``as shown in (Ladefoged {[}3{]})''.
 
-References are to be numbered in alphabetical order. Please
-double-check the final version of your paper with regard to the
-correct correspondence of references to their numbers.
+References are to be numbered in alphabetical order. Please double-check
+the final version of your paper with regard to the correct
+correspondence of references to their numbers.
 
@@
 hyperlinks to web pages are listed in the references section. If
-required, line breaks can be placed within URLs after slashes or
-dashes (cf.~\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that no hyphens
-are inserted.
+required, line breaks can be placed within URLs after slashes or dashes
+(cf.\textasciitilde{}\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that
+no hyphens are inserted.
 
@@
 
-Multimedia data that are part of the paper are to be embedded in the submitted PDF; they cannot be submitted as supplementary data.
-Any images are to be included in the paper as Figures (see Section 2.3 above).
-It is the authors' responsibility to check image quality ahead of submission.
-Audio examples are to be embedded within the PDF.
-To do this, authors can generate the PDF, and then embed the audio files using Adobe Acrobat Professional, or other software that offers the same outcome, so that the audio is included in the PDF.
-The presence of audio data should be identified in the text.
+Multimedia data that are part of the paper are to be embedded in the
+submitted PDF; they cannot be submitted as supplementary data. Any
+images are to be included in the paper as Figures (see Section 2.3
+above). It is the authors' responsibility to check image quality ahead
+of submission. Audio examples are to be embedded within the PDF. To do
+this, authors can generate the PDF, and then embed the audio files using
+Adobe Acrobat Professional, or other software that offers the same
+outcome, so that the audio is included in the PDF. The presence of audio
+data should be identified in the text.
 
-We encourage authors to illustrate video data using still photographs from the video, and to include them as figures in the PDF.
-We cannot accept embedded video files, but authors are welcome to refer readers to a URL on the internet where these can be viewed.
+We encourage authors to illustrate video data using still photographs
+from the video, and to include them as figures in the PDF. We cannot
+accept embedded video files, but authors are welcome to refer readers to
+a URL on the internet where these can be viewed.
 
@@
 
-In ICPhS 2023 submissions, an anonymous reviewing process will be used. This means that for the first submission the name(s) of the author(s) and
-their affiliation(s) \emph{must not} be mentioned.
-In addition, please refrain from using acknowledgements.
-Please also try to make your own previous research as anonymous as possible.
-As an example: do not write "In our previous study \cite{Stevens:1999} we could show\ldots" but "As shown in \cite{Stevens:1999}\ldots".
-Or refer to your own published or otherwise widely known work, and to that of the other authors, in the "Julius Caesar style", i.e.
-in the third person (for example: his work, her work, their work).
-Reference as "anonymous" only work that you or the other authors have submitted for publication, but which has not yet been published, e.g.
-\cite{Anonymous:submitted}.
+In ICPhS 2023 submissions, an anonymous reviewing process will be used.
+This means that for the first submission the name(s) of the author(s)
+and their affiliation(s) \emph{must not} be mentioned. In addition,
+please refrain from using acknowledgements. Please also try to make your
+own previous research as anonymous as possible. As an example: do not
+write ``In our previous study \cite{Stevens:1999} we could show\ldots''
+but ``As shown in \cite{Stevens:1999}\ldots''. Or refer to your own
+published or otherwise widely known work, and to that of the other
+authors, in the ``Julius Caesar style'', i.e.~in the third person (for
+example: his work, her work, their work). Reference as ``anonymous''
+only work that you or the other authors have submitted for publication,
+but which has not yet been published, e.g.~\cite{Anonymous:submitted}.
 
-Please make sure that no author details appear in the Document Properties of the PDF file.
+Please make sure that no author details appear in the Document
+Properties of the PDF file.
 \emph{For the revised paper submission author details are of course needed}.
@@
 
-References are formatted using the IEEE standard (available in various reference management systems like Zotero or Mendeley).
-If you do not use a reference management system, please use the examples provided for monographs \cite{Fant:1960}, contributions to volumes \cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles in conference proceedings \cite{Ladefoged:2003}.
-Abbreviations of
-well-known conferences and journals are possible \cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016} are referenced according to authors' instructions.
+References are formatted using the IEEE standard (available in various
+reference management systems like Zotero or Mendeley). If you do not use
+a reference management system, please use the examples provided for
+monographs \cite{Fant:1960}, contributions to volumes
+\cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles
+in conference proceedings \cite{Ladefoged:2003}. Abbreviations of
+well-known conferences and journals are possible
+\cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016}
+are referenced according to authors' instructions.
 
@@
 
-```{r, 'chunk_ex'}
-2 + 2
-```
+\begin{Shaded}
+\begin{Highlighting}[]
+\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
+\end{Highlighting}
+\end{Shaded}
 
-This includes inline code: 2 + 2 = `r 2 + 2`.
+\begin{verbatim}
+## [1] 4
+\end{verbatim}
+
+This includes inline code: 2 + 2 = 4.
 
@@
 
-```{r, 'load_script', echo=TRUE, message=FALSE, warning=FALSE}
-source("./includes/scripts/analysis.R")
-```
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{source}\NormalTok{(}\StringTok{"./includes/scripts/analysis.R"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-Which means you have access to any objects assigned in the script, like Figure 
-\ref{fig:ex_plot}:
+Which means you have access to any objects assigned in the script, like
+Figure \ref{fig:ex_plot}:
 
-```{r, 'ex_plot', fig.align='center', fig.width=2.75, fig.height=1.5, fig.cap.top="This is a figure caption.", fig.pos='h', fig.lp='fig:'} 
-cars_plot
-```
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cars\_plot}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[!h]
+\caption{This is a figure caption.}\label{fig:ex_plot}
+
+\begin{center}\includegraphics{main_files/figure-latex/ex_plot-1} \end{center}
+
+\end{figure}
 
@@
 \theendnotes
+
+\end{document}

### 2024-03-22 manuscript/main.Rmd

@@
   - \usepackage[T1]{fontenc}
+  - \usepackage{tipa}
   #- \usepackage[leqno,fleqn]{amsmath}
@@
 
-\begin{itemize}
-  \item they must be written in English
-  \item they must not contain the name(s) of the author(s) (for
-anonymous review)
-  \item the maximum is four pages for Congress papers, plus up to one additional page of references. Papers for Plenary lectures can be longer
-  \item they must be submitted in PDF format (cf.\ Section~4)
-  \item the paper submission will occur via a web interface
-\end{itemize}
-This paper template can be found on the conference website. If there
-are special questions or wishes regarding paper preparation and
-submission for ICPhS 2023, correspondence should be addressed to
-icphs2023@guarant.cz. Information for full paper submission will be
-available on the web at \url{https://www.icphs2023.org/}.
-
 \section{Background Literature}
-
-Write something?
 
@@
 279).
-In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [sɛl] - *zel* [zɛl] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019).
+In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019).
 Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022).
-For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tórəs] and *torres antigues* [torəzəntiɣəs].
+For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].
 
@@
 These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.
+
+\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}
+
+The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual (Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts (Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
+Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine.
+Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class (Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).
+
+Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers.
+Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older.
+Contrary, men who produce intervocalic [z] do not receive these negative evaluations.
+Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing.
+García (2020) hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties.
+Interestingly, the mean voicing rates of females are roughly two age groups behind those of males.
+Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.
+
+With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan (Davidson, 2014, 2015).
+According to Sinner (2002), intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness (Davidson, 2019; Sinner, 2002).
+In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish.
+Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not.
+Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization.
+These results are in line with previous findings on intervocalic [z] in Catalonian Spanish.
+For instance, Silverstein (2003) found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride.
+Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness (Davidson, 2019).
 

### 2024-03-22 manuscript/main.tex

@@
 \usepackage[T1]{fontenc}
+\usepackage{tipa}
 \ifLuaTeX
@@
 
-\begin{itemize}
-  \item they must be written in English
-  \item they must not contain the name(s) of the author(s) (for
-anonymous review)
-  \item the maximum is four pages for Congress papers, plus up to one additional page of references. Papers for Plenary lectures can be longer
-  \item they must be submitted in PDF format (cf.\ Section~4)
-  \item the paper submission will occur via a web interface
-\end{itemize}
-
-This paper template can be found on the conference website. If there are
-special questions or wishes regarding paper preparation and submission
-for ICPhS 2023, correspondence should be addressed to
-\href{mailto:icphs2023@guarant.cz}{\nolinkurl{icphs2023@guarant.cz}}.
-Information for full paper submission will be available on the web at
-\url{https://www.icphs2023.org/}.
-
 \section{Background Literature}
-
-Write something?
 
@@
 Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
-phonemically, which leads to minimal pairs like \emph{cel} {[}sɛl{]} -
-\emph{zel} {[}zɛl{]} and \emph{fossa} {[}fósa{]} - \emph{fosa}
-{[}fóza{]} (Carrera-Sabaté \& Prieto, 2012-2019). Conversely, at the end
-of the word, this contrast is neutralized, being the voicing or
-devoicing of the alveolar fricative dependent on the following sound
-(Institut d'Estudis Catalans, 2022). For example, if it is followed by a
-vowel, the alveolar fricative is voiced: \emph{torres} {[}tórəs{]} and
-\emph{torres antigues} {[}torəzəntiɣəs{]}.
+phonemically, which leads to minimal pairs like \emph{cel}
+{[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
+\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} (Carrera-Sabaté \&
+Prieto, 2012-2019). Conversely, at the end of the word, this contrast is
+neutralized, being the voicing or devoicing of the alveolar fricative
+dependent on the following sound (Institut d'Estudis Catalans, 2022).
+For example, if it is followed by a vowel, the alveolar fricative is
+voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
+{[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.
 
@@
 fricative /z/ in Catalan.
+
+\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}
+
+The social meaning and the perception of intervocalic {[}z{]} in Spanish
+has been studied in a variety of contexts, including both monolingual
+(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \& García,
+2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts
+(Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been
+categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et
+al., 2023; García, 2015)With respect to the former, it has been found
+that for listeners from San José, Costa Rica, the covert attitudes
+associated with intervocalic {[}z{]} vary significantly by speaker
+gender. Consequently, male speakers who use {[}z{]} tend to be
+positively evaluated, and generally perceived as nice, local, confident,
+and masculine. Conversely, female speakers who use {[}z{]} are only
+afforded negative associations of lower education and social class
+(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).
+
+Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is
+evaluated differently in male and female speakers. Accordingly, women
+who produce intervocalic {[}z{]} are viewed as significantly lower
+status, less pleasant, and older. Contrary, men who produce intervocalic
+{[}z{]} do not receive these negative evaluations. Additionally, this
+variable has been found to serve as a marker of regional identity, with
+voicing being associated with neighboring Highland Ecuadorian dialects
+(Cuenca and Quito) that exhibit systematic voicing. García (2020)
+hypothesizes that young men are the leaders of this change, essentially
+due to the fact that they are in greater contact with these highland
+varieties. Interestingly, the mean voicing rates of females are roughly
+two age groups behind those of males. Consequently, it is believed that
+it is not necessarily the case that women in Ecuador are avoiding
+voicing like San José women due to its negative evaluations, but rather
+that they are further behind in the adoption of the incoming variant.
+
+With respect to the social meaning and the perception of intervocalic
+{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
+in Catalan-Spanish contact situations, given its robust usage by
+Catalan-Spanish bilinguals and, in particular, by those with greatest
+exposure to and usage of Catalan (Davidson, 2014, 2015). According to
+Sinner (2002), intervocalic {[}z{]} can be considered a possible
+linguistic marker of Catalonian Spanish, despite its relative lack of
+social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner
+(2002) interviewed 12 speakers of Barcelonan Spanish and monolingual
+Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of
+linguistic features of Catalanized Spanish. Results showed that only
+select Barcelona speakers perceived intervocalic {[}z{]} as a linguistic
+marker of Catalonian Spanish, whereas monolingual Spanish-speakers from
+Madrid did not. Additionally, Davidson (2019) found that, in Barcelona,
+the indexical field of intervocalic {[}z{]} features exclusively
+positive associations with bilingualism and local solidarity, which
+warrants its gradual adoption in the speech community, unlike other
+phonetic markers of Catalonian Spanish that are the topic of overt
+social ridicule, such as lateral velarization. These results are in line
+with previous findings on intervocalic {[}z{]} in Catalonian Spanish.
+For instance, Silverstein (2003) found that intervocalic {[}z{]} indexes
+a local Barcelona identity, and the variant then acquires positive
+associations with bilingualism and cultural pride. Overall, these
+findings suggest that intervocalic {[}z{]} likely functions as a
+sociolinguistic indicator of Catalonian Spanish within the Catalonian
+speech community, indexing Catalan-Spanish bilingual membership while
+lacking any considerable degree of generalized conscious awareness
+(Davidson, 2019).
 

### 2024-04-23 scripts/models.R
models + plots

@@ -1,16 +1,186 @@
 
 # Load libraries
-# library("tidyverse")
 library("here")
 library("lme4")
-# library("dplyr")
 
 
 # Load data
-subset_pbc <- read.csv(here("data", "subsets", "subset_pbc.csv"))
-subset_accent <- read.csv(here("data", "subsets", "subset_accent.csv"))
+subset_estudis <- read.csv(here("data", "subsets", "subset_estudis.csv"))
 subset_classe <- read.csv(here("data", "subsets", "subset_classe.csv"))
+subset_intelligent <- read.csv(here("data", "subsets", "subset_intelligent.csv"))
+subset_simpatic <- read.csv(here("data", "subsets", "subset_simpatic.csv"))
+subset_amable <- read.csv(here("data", "subsets", "subset_amable.csv"))
+subset_fiar <- read.csv(here("data", "subsets", "subset_fiar.csv"))
+subset_ciutat <- read.csv(here("data", "subsets", "subset_ciutat.csv"))
+subset_pbc <- read.csv(here("data", "subsets", "subset_pbc.csv"))
 subset_oc <- read.csv(here("data", "subsets", "subset_oc.csv"))
+subset_accent <- read.csv(here("data", "subsets", "subset_accent.csv"))
+subset_agradable <- read.csv(here("data", "subsets", "subset_agradable.csv"))
+
+
+# Build models (té estudis, 1)
+mod_0_estudis <- lm(value ~ 1, data = subset_estudis)
+
+mod_1_estudis <- lm(value ~ 1 + phoneme, data = subset_estudis)
+
+mod_2_estudis <- lm(value ~ 1 + phoneme + mother_tongue, data = subset_estudis)
+
+mod_3_estudis <- lm(value ~ 1 + phoneme + mother_tongue + province, 
+                    data = subset_estudis)
+
+mod_4_estudis <- lm(value ~ 1 + phoneme + mother_tongue + province + 
+                      phoneme:mother_tongue:province, data = subset_estudis)
+
+summary(mod_0_estudis)
+summary(mod_1_estudis)
+summary(mod_2_estudis)
+summary(mod_3_estudis)
+summary(mod_4_estudis)
+
+anova(mod_0_estudis, mod_1_estudis, mod_2_estudis, mod_3_estudis, mod_4_estudis)
+
+
+
[…]

### 2024-04-23 manuscript/main.Rmd

@@
 
-Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015).
-In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p.
-279).
-In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019).
-Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022).
+Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}.
+In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts **(Lloret, 2011; Recasens Vives, 1996, p.
+279)**.
+In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] **(Carrera-Sabaté & Prieto, 2012-2019)**.
+Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound **(Institut d’Estudis Catalans, 2022)**.
 For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].
 
-Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process (Hualde & Colina, 2015).
-Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) (Hualde & Colina, 2015; Hualde et al., 2020).
+Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process \cite{hualde_colina_2015}.
+Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) **(Hualde & Colina, 2015; Hualde et al., 2020)**.
 Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process.
-It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial (Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020).
+It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.
 
-There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules (Institut d'Estudis Catalans, 2022), the [z] should be produced, but [s] is produced instead.
-This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas (Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
-268).
+There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead.
+This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
+268)**.
 The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia.
-Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia (Davidson, 2020).
+Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.
 
@@
 
-The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual (Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts (Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
+The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual **(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003)** and bilingual contexts **(Davidson, 2014, 2015, 2019; Sinner, 2002)**, despite having been categorized as a low-salient variant **(Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)**.
+With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
 Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine.
-Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class (Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).
+Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class **(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003)**.
 
-Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers.
+Similarly, **García (2019)** shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers.
 Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older.
@@
 Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing.
-García (2020) hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties.
+**García (2020)** hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties.
 Interestingly, the mean voicing rates of females are roughly two age groups behind those of males.
@@
 
-With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan (Davidson, 2014, 2015).
-According to Sinner (2002), intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness (Davidson, 2019; Sinner, 2002).
+With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan **(Davidson, 2014, 2015)**.
+According to **Sinner (2002)**, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness **(Davidson, 2019; Sinner, 2002)**.
 In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish.
@@
 These results are in line with previous findings on intervocalic [z] in Catalonian Spanish.
-For instance, Silverstein (2003) found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride.
-Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness (Davidson, 2019).
+For instance, **Silverstein (2003)** found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride.
+Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness **(Davidson, 2019)**.
 

### 2024-04-23 manuscript/main.tex

@@
 Both Catalan and Peninsular Spanish sound inventories have an phoneme
-/s/, which can be classified as a voiceless alveolar fricative (Hualde
-\& Colina, 2015). In addition, most of the Catalan varieties have a
-voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens
-Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
-phonemically, which leads to minimal pairs like \emph{cel}
+/s/, which can be classified as a voiceless alveolar fricative
+\cite{hualde_colina_2015}. In addition, most of the Catalan varieties
+have a voiced alveolar phoneme in intervocalic contexts \textbf{(Lloret,
+2011; Recasens Vives, 1996, p.~279)}. In onset position, /s/ and /z/
+contrast phonemically, which leads to minimal pairs like \emph{cel}
 {[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
-\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} (Carrera-Sabaté \&
-Prieto, 2012-2019). Conversely, at the end of the word, this contrast is
-neutralized, being the voicing or devoicing of the alveolar fricative
-dependent on the following sound (Institut d'Estudis Catalans, 2022).
-For example, if it is followed by a vowel, the alveolar fricative is
-voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
+\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} \textbf{(Carrera-Sabaté
+\& Prieto, 2012-2019)}. Conversely, at the end of the word, this
+contrast is neutralized, being the voicing or devoicing of the alveolar
+fricative dependent on the following sound \textbf{(Institut d'Estudis
+Catalans, 2022)}. For example, if it is followed by a vowel, the
+alveolar fricative is voiced: \emph{torres} {[}tór\textschwa{}s{]} and
+\emph{torres antigues}
 {[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.

### 2024-04-25 manuscript/main.Rmd
update

@@
 Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness **(Davidson, 2019)**.
+
+\section{Methods}
+
+\subsection{Materials}
+
+Data were collected using an online survey hosted on Qualtrics that consisted of two parts.
+The first part, a demographic questionnaire, collected information regarding participant age, gender, place of origin, place of residence, mother tongue, family income, and level of studies.
+The second portion of the survey consisted of a perception task where participants were presented with a series of stimuli containing the target phoneme contrast (intervocalic voiced and voiceless alveolar fricative) followed by 11 evaluative statements related to social characteristics such as level of studies, friendliness, social class, and ethnic origin.
+These evaluative statements were then followed by a 100-point draggable sliding scale with which participants indicated their level of agreement or disagreement with the content of the statement, where 0 was “totally disagree” and 100 was “totally agree”.
+
+The target stimuli were recorded by a 28-year-old female native speaker of Catalan in a whisper room 6084 E sound booth using a personal communication device.
+The stimuli were between 3 and 4 seconds long and were controlled for utterance length (5 words), syllable count (2 syllables for the target word and 3 for the following adjective), and lexical stress (target sound in unstressed position).
+A total of 12 target stimuli containing intervocalic voiced /s/ were recorded, as well as different samples of the voiceless /s/.
+The target sentences were manipulated in Praat by splicing the voiced /s/ and pasting the voiceless /s/ in order to create 2 conditions: 12 stimuli containing intervocalic voiced /s/ and 12 stimuli containing intervocalic voiceless /s/, for a total of 24 experimental sentences.
+These sentences were combined with 32 fillers targeting other phonemic contrasts, for a total of 56 sentences randomly distributed in two different versions.
+Each included 12 experimental stimuli (6 x 2 conditions) and 16 fillers (8 x 2 conditions).
+
+\subsection{Procedure}
+
+Participants self-identified for this study.
+Recruitment materials in Catalan with information about the study, the requirements to participate and a link to the survey were posted on different social networks and online communication channels, such as WhatsApp and Instagram.
+Participants who were interested and wanted to participate clicked on the link and read the study description.
+Those who decided to participate completed the survey on their own electronic devices and at their own pace without being observed.
+On average, the survey took approximately 15 minutes to be completed.
+The survey followed a between-subjects design.
+Therefore, participants were randomly assigned to one of the two versions.
+
+\subsection{Participants}
+
+A total of 78 responses to the questionnaire were collected.
+Of these, 56 responses were incomplete and were therefore discarded.
+Data are drawn from a total of 22 participants (13 females) aged between 18 and 72 years old (mean = 38) from Barcelona and Girona.
+In order to be eligible to take part in the study, participants had to meet the following three requirements: to be over 18 years of age, to be from Catalonia, and to be proficient in both Catalan and Spanish.
+
+\subsection{Data analysis}
+
+Data were analyzed in R (R Core Team, 2023).
+First, we performed a descriptive analysis that consisted of creating a series of plots and tables to summarize and describe the features of the dataset and to identify potential trends.
+After that, we performed inferential statistical analysis.
+For each of the 11 evaluative statements, we created a series of hierarchically nested linear regression models using the lme4 package (Bates et al., 2015) with attitudinal sliding scale responses as the dependent variable, and phoneme, mother tongue, and province as potential predictors.
+To determine the best fit model for each statement, we used the anova() function (Fox & Weisberg, 2019).
 

### 2024-04-25 manuscript/main.tex

@@
 distinguished between /s/ and /z/ in intervocalic position within a
-word, although these sounds merged through a devoicing process (Hualde
-\& Colina, 2015). Nevertheless, in those modern varieties where the /s/
-in coda position is produced, such as in Peninsular Spanish, this
-voiceless phoneme might assimilate to the following sound, especially if
-it is a voiced consonant (\emph{mismo}, {[}mízmo{]}), rather than a
-vowel (\emph{las alas}, {[}lasálas{]}) (Hualde \& Colina, 2015; Hualde
-et al., 2020). Hence, Spanish inventory does not have the phoneme /z/,
-but {[}z{]} may occur as an allophone of /s/, due to a voice
-assimilation process. It is important to mention, though, that it has
-been claimed that this voicing is generally not complete, but partial
-(Benet et al., 2012; Hualde \& Colina, 2015; Hualde et al., 2020).
+word, although these sounds merged through a devoicing process
+\cite{hualde_colina_2015}. Nevertheless, in those modern varieties where
+the /s/ in coda position is produced, such as in Peninsular Spanish,
+this voiceless phoneme might assimilate to the following sound,
+especially if it is a voiced consonant (\emph{mismo}, {[}mízmo{]}),
+rather than a vowel (\emph{las alas}, {[}lasálas{]}) \textbf{(Hualde \&
+Colina, 2015; Hualde et al., 2020)}. Hence, Spanish inventory does not
+have the phoneme /z/, but {[}z{]} may occur as an allophone of /s/, due
+to a voice assimilation process. It is important to mention, though,
+that it has been claimed that this voicing is generally not complete,
+but partial \textbf{(Benet et al., 2012; Hualde \& Colina, 2015; Hualde
+et al., 2020)}.
 
@@
 the Catalan alveolar fricative, in contexts where, according to
-prescriptivist rules (Institut d'Estudis Catalans, 2022), the {[}z{]}
-should be produced, but {[}s{]} is produced instead. This phenomenon has
-mainly been attested in those areas where there has been an extended
-contact between Catalan and Spanish, such asthe so-called ``franja de
-Aragón'' (borderland between Catalonia and Aragon), some areas of the
-Valencian Community, and Barcelona and other urban areas (Benet et al.,
-2012; Davidson, 2020; Recasens Vives, 1996, p.~268). The opposite
-pattern has also been observed - that is, the voicing of the alveolar
-fricative in the Spanish spoken in Barcelona and Valencia.
-Interestingly, although in both Barcelona and Valencia Catalan and
-Spanish coexist and influence each other, it seems that Catalan has a
-greater influence on Spanish in Barcelona, but Spanish has a greater
-influence on Catalan in Valencia (Davidson, 2020).
+prescriptivist rules \textbf{(Institut d'Estudis Catalans, 2022)}, the
+{[}z{]} should be produced, but {[}s{]} is produced instead. This
+phenomenon has mainly been attested in those areas where there has been
+an extended contact between Catalan and Spanish, such asthe so-called
+``franja de Aragón'' (borderland between Catalonia and Aragon), some
+areas of the Valencian Community, and Barcelona and other urban areas
+\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
+p.~268)}. The opposite pattern has also been observed - that is, the
+voicing of the alveolar fricative in the Spanish spoken in Barcelona and
+Valencia. Interestingly, although in both Barcelona and Valencia Catalan
+and Spanish coexist and influence each other, it seems that Catalan has
+a greater influence on Spanish in Barcelona, but Spanish has a greater
+influence on Catalan in Valencia \textbf{(Davidson, 2020)}.
 
@@
 has been studied in a variety of contexts, including both monolingual
-(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \& García,
-2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts
-(Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been
-categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et
-al., 2023; García, 2015)With respect to the former, it has been found
-that for listeners from San José, Costa Rica, the covert attitudes
-associated with intervocalic {[}z{]} vary significantly by speaker
-gender. Consequently, male speakers who use {[}z{]} tend to be
-positively evaluated, and generally perceived as nice, local, confident,
-and masculine. Conversely, female speakers who use {[}z{]} are only
-afforded negative associations of lower education and social class
-(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).
+\textbf{(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \&
+García, 2017; García, 2015, 2019; Silverstein, 2003)} and bilingual
+contexts \textbf{(Davidson, 2014, 2015, 2019; Sinner, 2002)}, despite
+having been categorized as a low-salient variant \textbf{(Chappell,
+2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
+former, it has been found that for listeners from San José, Costa Rica,
+the covert attitudes associated with intervocalic {[}z{]} vary
+significantly by speaker gender. Consequently, male speakers who use
+{[}z{]} tend to be positively evaluated, and generally perceived as
+nice, local, confident, and masculine. Conversely, female speakers who
+use {[}z{]} are only afforded negative associations of lower education
+and social class \textbf{(Chappell, 2016; Chappell et al., 2023;
+Silverstein, 2003)}.
 
-Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is
-evaluated differently in male and female speakers. Accordingly, women
-who produce intervocalic {[}z{]} are viewed as significantly lower
+Similarly, \textbf{García (2019)} shows that, in Ecuador, intervocalic
+/s/ is evaluated differently in male and female speakers. Accordingly,
+women who produce intervocalic {[}z{]} are viewed as significantly lower
 status, less pleasant, and older. Contrary, men who produce intervocalic
@@
 voicing being associated with neighboring Highland Ecuadorian dialects
-(Cuenca and Quito) that exhibit systematic voicing. García (2020)
-hypothesizes that young men are the leaders of this change, essentially
-due to the fact that they are in greater contact with these highland
-varieties. Interestingly, the mean voicing rates of females are roughly
-two age groups behind those of males. Consequently, it is believed that
-it is not necessarily the case that women in Ecuador are avoiding
-voicing like San José women due to its negative evaluations, but rather
-that they are further behind in the adoption of the incoming variant.
+(Cuenca and Quito) that exhibit systematic voicing. \textbf{García
+(2020)} hypothesizes that young men are the leaders of this change,
+essentially due to the fact that they are in greater contact with these
+highland varieties. Interestingly, the mean voicing rates of females are
+roughly two age groups behind those of males. Consequently, it is
+believed that it is not necessarily the case that women in Ecuador are
+avoiding voicing like San José women due to its negative evaluations,
+but rather that they are further behind in the adoption of the incoming
+variant.
 
@@
 Catalan-Spanish bilinguals and, in particular, by those with greatest
-exposure to and usage of Catalan (Davidson, 2014, 2015). According to
-Sinner (2002), intervocalic {[}z{]} can be considered a possible
-linguistic marker of Catalonian Spanish, despite its relative lack of
-social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner
-(2002) interviewed 12 speakers of Barcelonan Spanish and monolingual
-Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of
-linguistic features of Catalanized Spanish. Results showed that only
-select Barcelona speakers perceived intervocalic {[}z{]} as a linguistic
-marker of Catalonian Spanish, whereas monolingual Spanish-speakers from
-Madrid did not. Additionally, Davidson (2019) found that, in Barcelona,
-the indexical field of intervocalic {[}z{]} features exclusively
-positive associations with bilingualism and local solidarity, which
-warrants its gradual adoption in the speech community, unlike other
-phonetic markers of Catalonian Spanish that are the topic of overt
-social ridicule, such as lateral velarization. These results are in line
-with previous findings on intervocalic {[}z{]} in Catalonian Spanish.
-For instance, Silverstein (2003) found that intervocalic {[}z{]} indexes
-a local Barcelona identity, and the variant then acquires positive
+exposure to and usage of Catalan \textbf{(Davidson, 2014, 2015)}.
+According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
+considered a possible linguistic marker of Catalonian Spanish, despite
+its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
+2002)}. In their study, Sinner (2002) interviewed 12 speakers of
+Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27
+to 41 regarding their awareness of linguistic features of Catalanized
+Spanish. Results showed that only select Barcelona speakers perceived
+intervocalic {[}z{]} as a linguistic marker of Catalonian Spanish,
+whereas monolingual Spanish-speakers from Madrid did not. Additionally,
+Davidson (2019) found that, in Barcelona, the indexical field of
+intervocalic {[}z{]} features exclusively positive associations with
+bilingualism and local solidarity, which warrants its gradual adoption
+in the speech community, unlike other phonetic markers of Catalonian
+Spanish that are the topic of overt social ridicule, such as lateral
+velarization. These results are in line with previous findings on
+intervocalic {[}z{]} in Catalonian Spanish. For instance,
+\textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
+local Barcelona identity, and the variant then acquires positive
 associations with bilingualism and cultural pride. Overall, these
@@
 lacking any considerable degree of generalized conscious awareness
-(Davidson, 2019).
+\textbf{(Davidson, 2019)}.
+
+\section{Methods}
+
+\subsection{Materials}
+
+Data were collected using an online survey hosted on Qualtrics that
+consisted of two parts. The first part, a demographic questionnaire,
+collected information regarding participant age, gender, place of
+origin, place of residence, mother tongue, family income, and level of
+studies. The second portion of the survey consisted of a perception task
+where participants were presented with a series of stimuli containing
+the target phoneme contrast (intervocalic voiced and voiceless alveolar
+fricative) followed by 11 evaluative statements related to social
+characteristics such as level of studies, friendliness, social class,
+and ethnic origin. These evaluative statements were then followed by a
+100-point draggable sliding scale with which participants indicated
+their level of agreement or disagreement with the content of the
+statement, where 0 was ``totally disagree'' and 100 was ``totally
+agree''.
+
+The target stimuli were recorded by a 28-year-old female native speaker
+of Catalan in a whisper room 6084 E sound booth using a personal
+communication device. The stimuli were between 3 and 4 seconds long and
+were controlled for utterance length (5 words), syllable count (2
+syllables for the target word and 3 for the following adjective), and
+lexical stress (target sound in unstressed position). A total of 12
+target stimuli containing intervocalic voiced /s/ were recorded, as well
+as different samples of the voiceless /s/. The target sentences were
+manipulated in Praat by splicing the voiced /s/ and pasting the
+voiceless /s/ in order to create 2 conditions: 12 stimuli containing
+intervocalic voiced /s/ and 12 stimuli containing intervocalic voiceless
+/s/, for a total of 24 experimental sentences. These sentences were
+combined with 32 fillers targeting other phonemic contrasts, for a total
+of 56 sentences randomly distributed in two different versions. Each
+included 12 experimental stimuli (6 x 2 conditions) and 16 fillers (8 x
+2 conditions).
+
+\subsection{Procedure}
+
+Participants self-identified for this study. Recruitment materials in
+Catalan with information about the study, the requirements to
+participate and a link to the survey were posted on different social
+networks and online communication channels, such as WhatsApp and
+Instagram. Participants who were interested and wanted to participate
+clicked on the link and read the study description. Those who decided to
+participate completed the survey on their own electronic devices and at
+their own pace without being observed. On average, the survey took
+approximately 15 minutes to be completed. The survey followed a
+between-subjects design. Therefore, participants were randomly assigned
+to one of the two versions.
+
+\subsection{Participants}
+
+A total of 78 responses to the questionnaire were collected. Of these,
+56 responses were incomplete and were therefore discarded. Data are
+drawn from a total of 22 participants (13 females) aged between 18 and
+72 years old (mean = 38) from Barcelona and Girona. In order to be
+eligible to take part in the study, participants had to meet the
+following three requirements: to be over 18 years of age, to be from
+Catalonia, and to be proficient in both Catalan and Spanish.
+
+\subsection{Data analysis}
+
+Data were analyzed in R (R Core Team, 2023). First, we performed a
+descriptive analysis that consisted of creating a series of plots and
+tables to summarize and describe the features of the dataset and to
+identify potential trends. After that, we performed inferential
+statistical analysis. For each of the 11 evaluative statements, we
+created a series of hierarchically nested linear regression models using
+the lme4 package (Bates et al., 2015) with attitudinal sliding scale
+responses as the dependent variable, and phoneme, mother tongue, and
+province as potential predictors. To determine the best fit model for
+each statement, we used the anova() function (Fox \& Weisberg, 2019).
 

### 2024-04-25 manuscript/main.Rmd
manuscript

@@
 
+\subsection{The present study}
+**INTRO**. Accordingly, the research questions and hypotheses are as follows:
+1. What covert attitudes are linked to the perception Catalan intervocalic /s/ as either [s] or [z]?
+2. What is the relationship between  individual social factors (mother tongue, province of origin) and covert attitudes towards Catalan intervocalic /s/?
+
 \section{Methods}
@@
 To determine the best fit model for each statement, we used the anova() function (Fox & Weisberg, 2019).
+
+\section{Results}
+The statistical analysis did not reveal a significant effect of phoneme on the evaluation of the speaker (p > .05) for any of the 11 evaluative statements.
+However, the model output did show a significant association between mother tongue and province for some of the statements.
+In what follows, we will explore the results for those statements.
+
+Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' showed similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}.
+Those who have both Catalan and Spanish as their L1 rated the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/pbc.png}
+\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/oc.png}
+\caption{Ratings of "És d'origen català" as a function of mother tongue}\label{fig:oc}
+\end{center}
+\end{figure}
+
+Participant responses to the statements *És simpàtic* 'S/he is nice' and *És de fiar* 'S/he is trustworthy' showed similar results, as illustrated in Figures \ref{fig:simpatic} and {fig:fiar}.
+Those who are from and live in the province of Barcelona give higher ratings to the speaker, compared to those who are from and live in Girona.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/simpatic.png}
+\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/fiar.png}
+\caption{Ratings of "És de fiar" as a function of mother tongue}\label{fig:fiar}
+\end{center}
+\end{figure}
+
+\section{Discussion and Conclusions}
+The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a Catalan native speaker.
+The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.
+
+Contrary to what we expected, phoneme was not a significant predictor.
+One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] or, if they could, this difference in production was not changing their attitudes towards the speaker.
+This finding is interesting considering that, in our stimuli, the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
+268)**, and, from a prescriptivist point of view, it should be avoided **(Institut d'Estudis Catalans, 2022)**.
+Despite the lack of significance of this predictor, it should be mentioned that the speaker received more positive ratings when the [z] was being produced.
+Although at this point we can only hypothesize, it could be that the combination of the phoneme /s/ with the precedent vowel, might give the listener more cues and have a stronger social meaning.
+Future studies should address this issue in order to explore if the combination of vowel + /s/ carries other social connotations.
+**REVISAR**
+
+**CONTINUAR**
 

### 2024-04-25 scripts/plots.R

@@ -29,9 +29,11 @@ subset_pbc %>%
   scale_x_discrete(labels = c("Spanish", "Catalan", "Both")) +
   theme(legend.key.size = unit(1, "lines"),
     legend.key.width = unit(1.5, "lines"),
-    legend.spacing.x = unit(0.5, "lines")) +
-  labs(title = "Statement: Parla bé català") +
+    legend.spacing.x = unit(0.5, "lines"),
+    legend.position = "bottom") +
+  #labs(title = "Statement: Parla bé català") +
   guides(fill = guide_legend(override.aes = list(size = 0.5)))
+ggsave(here("manuscript", "includes", "figures", "pbc.png"))
 
 # Sembla que la gent no distingueix entre [s] i [z] o, si ho fa, els hi és igual
 # En general, els cat_cast valoren més positivament a la persona que escolten,
@@ -56,9 +58,11 @@ subset_oc %>%
   scale_x_discrete(labels = c("Spanish", "Catalan", "Both")) +
   theme(legend.key.size = unit(1, "lines"),
         legend.key.width = unit(1.5, "lines"),
-        legend.spacing.x = unit(0.5, "lines"))+
-  labs(title = "Statement: És d'origen català") +
+        legend.spacing.x = unit(0.5, "lines"),
+        legend.position = "bottom")+
+  #labs(title = "Statement: És d'origen català") +
   guides(fill = guide_legend(override.aes = list(size = 0.5)))
+ggsave(here("manuscript", "includes", "figures", "oc.png"))
 
 # És pràcticament igual a l'anterior
 
@@ -104,9 +108,11 @@ subset_simpatic %>%
   scale_fill_viridis_d(end = 0.9, labels = c("Barcelona", "Girona")) +
   theme(legend.key.size = unit(1, "lines"),
         legend.key.width = unit(1.5, "lines"),
-        legend.spacing.x = unit(0.5, "lines"))+
-  labs(title = "Statement: És simpàtic") +
+        legend.spacing.x = unit(0.5, "lines"),
+        legend.position = "bottom")+
+  #labs(title = "Statement: És simpàtic") +
   guides(fill = guide_legend(override.aes = list(size = 0.5)))
+ggsave(here("manuscript", "includes", "figures", "simpatic.png"))
 
 
 
@@ -136,9 +142,11 @@ subset_fiar %>%
   scale_fill_viridis_d(end = 0.9, labels = c("Barcelona", "Girona")) +
[…]

### 2024-04-25 manuscript/main.tex

@@
 
+\subsection{The present study}
+
+\textbf{INTRO}. Accordingly, the research questions and hypotheses are
+as follows: 1. What covert attitudes are linked to the perception
+Catalan intervocalic /s/ as either {[}s{]} or {[}z{]}? 2. What is the
+relationship between individual social factors (mother tongue, province
+of origin) and covert attitudes towards Catalan intervocalic /s/?
+
 \section{Methods}
@@
 each statement, we used the anova() function (Fox \& Weisberg, 2019).
+
+\section{Results}
+
+The statistical analysis did not reveal a significant effect of phoneme
+on the evaluation of the speaker (p \textgreater{} .05) for any of the
+11 evaluative statements. However, the model output did show a
+significant association between mother tongue and province for some of
+the statements. In what follows, we will explore the results for those
+statements.
+
+Participant responses to the statements \emph{Parla bé català} `S/he
+speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
+origin' showed similar results, as illustrated in Figures \ref{fig:pbc}
+and \{fig:oc\}. Those who have both Catalan and Spanish as their L1
+rated the speaker more positively regardless of phoneme, compared to
+those whose L1 is either Catalan or Spanish.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/pbc.png}
+\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/oc.png}
+\caption{Ratings of "És d'origen català" as a function of mother tongue}\label{fig:oc}
+\end{center}
+\end{figure}
+
+Participant responses to the statements \emph{És simpàtic} `S/he is
+nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
+results, as illustrated in Figures \ref{fig:simpatic} and \{fig:fiar\}.
+Those who are from and live in the province of Barcelona give higher
+ratings to the speaker, compared to those who are from and live in
+Girona.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/simpatic.png}
+\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[height=15cm]{./includes/figures/fiar.png}
+\caption{Ratings of "És de fiar" as a function of mother tongue}\label{fig:fiar}
+\end{center}
+\end{figure}
+
+\section{Discussion and Conclusions}
+
+The aim of this exploration study was to investigate whether phoneme
+production ({[}s{]} vs {[}z{]}), mother tongue (Catalan vs Spanish vs
+both), and province of origin (Girona vs Barcelona) influenced the
+attitudes of the participants towards a Catalan native speaker. The
+results suggest that these predictors have a different effect on the
+ratings of the speaker depending on the social characteristic under
+evaluation.
+
+Contrary to what we expected, phoneme was not a significant predictor.
+One explanation for this divergent pattern could be that either
+participants could not distinguish between {[}s{]} and {[}z{]} or, if
+they could, this difference in production was not changing their
+attitudes towards the speaker. This finding is interesting considering
+that, in our stimuli, the production of the voiceless alveolar fricative
+in intervocalic position has been considered a result of Catalan in
+contact with Spanish \textbf{(Benet et al., 2012; Davidson, 2020;
+Recasens Vives, 1996, p.~268)}, and, from a prescriptivist point of
+view, it should be avoided \textbf{(Institut d'Estudis Catalans, 2022)}.
+Despite the lack of significance of this predictor, it should be
+mentioned that the speaker received more positive ratings when the
+{[}z{]} was being produced. Although at this point we can only
+hypothesize, it could be that the combination of the phoneme /s/ with
+the precedent vowel, might give the listener more cues and have a
+stronger social meaning. Future studies should address this issue in
+order to explore if the combination of vowel + /s/ carries other social
+connotations. \textbf{REVISAR}
 

### 2024-04-27 manuscript/main.Rmd
update

@@
 
-Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}.
+Both Catalan and Peninsular Spanish sound inventories have a phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}.
 In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts **(Lloret, 2011; Recasens Vives, 1996, p.

### 2024-04-27 manuscript/main.Rmd

@@
 
-Write something here
+Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean cost **(Hualde, 1992)**.
+It is safe to assume that there is a bidirectional influence between these languages: in that region, Catalan has influenced Spanish and Spanish has influenced Catalan in all linguistic levels **(Davidson, 2014, 2015)**.
+A crucial area of cross-linguistic influence might be the production of sounds, especially in those sounds who have different distribution in Catalan and Spanish.
+For example, previous studies have found that, in Barcelona, the production of Catalan vowels is different based on the L1 of the speaker, being it less target-like in L1 Spanish speakers compared to L1 Catalan speakers **(CITAR SEBASTIAN GALLES, etc)**.
+
+As for how Catalan people perceive these differences in production and, thus, their attitudes towards a less Catalan-like production of a given phoneme little is known.
+The aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test **(Lambert, 1967)**.
+The alveolar fricative was chosen because it exists in both Catalan and Spanish inventories, but their contextual distribution is different **(Llisterri, 2023)**.
+
+**PONER ALGO MÁS**
 
@@
 There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead.
-This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
+This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such as the so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
 268)**.
@@
 \subsection{The present study}
-**INTRO**. Accordingly, the research questions and hypotheses are as follows:
+**INTRO**.
+
+Accordingly, the research questions and hypotheses are as follows:
 1. What covert attitudes are linked to the perception Catalan intervocalic /s/ as either [s] or [z]?
@@
 
-Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' showed similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}.
-Those who have both Catalan and Spanish as their L1 rated the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish.
+Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' show similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}.
+Those who have both Catalan and Spanish as their L1 rate the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish.
+In addition, although it is not statistically significant, participants give higher ratings to the speaker when they are listening to those statements with [z].
 
@@
 \section{Discussion and Conclusions}
-The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a Catalan native speaker.
+The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a native speaker of Central Catalan.
 The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.
@@
 Contrary to what we expected, phoneme was not a significant predictor.
-One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] or, if they could, this difference in production was not changing their attitudes towards the speaker.
-This finding is interesting considering that, in our stimuli, the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
+One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] **(Davidson, 2019)** or, if they could, this difference in production was not changing their attitudes towards the speaker **(Sinner, 2002)**.
+This finding is interesting considering that the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
 268)**, and, from a prescriptivist point of view, it should be avoided **(Institut d'Estudis Catalans, 2022)**.
 Despite the lack of significance of this predictor, it should be mentioned that the speaker received more positive ratings when the [z] was being produced.
-Although at this point we can only hypothesize, it could be that the combination of the phoneme /s/ with the precedent vowel, might give the listener more cues and have a stronger social meaning.
-Future studies should address this issue in order to explore if the combination of vowel + /s/ carries other social connotations.
-**REVISAR**
+The fact that phoneme was not an important factor contrasts with **Davidson (2020)**, who analyzed covert and overt attitudes towards Catalonian Spanish and found that [z] was a significant predictor: it was considered a sociolinguistic indicator of Catalan-Spanish bilingual membership.
+Although at this point we can only hypothesize, it could be that [z] is perceptually more salient in Spanish than in Catalan, as the Spanish inventory does not have the phoneme /z/, but voicing of /s/ might partially occur **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.
+Another explanation could be that the /s/ alone does not provide enough social information; rather, the combination of the phoneme /s/ with the precedent vowel might give the listener more cues and have a stronger social meaning.
+Future studies should address this issue in order to explore if the combination of vowel + /s/ carries other social connotations than just the /s/ alone.
 
-**CONTINUAR**
+We expected mother tongue to be a strong predictor of the attitudes towards the production of intervocalic /s/, specifically, that L1 Catalan and bilingual speakers would give higher rates to those statements with [z].
+This hypothesis was only partially confirmed.
+As reported above, phoneme was not a factor influencing the attitudes, and mother tongue was only a significant predictor in the two statements related to Catalan identity, namely *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin'.
+Interestingly, bilinguals gave higher ratings to the speaker, followed by L1 Catalan participants and L1 Spanish participants.
+A potential explanation for this finding, that should be verified in future research, could be that L1 Catalan speakers have specific ideas (higher standards) towards the 'proper' Catalan.
+In addition, some studies have suggested that those Catalan people who grow up in Spanish-speaking households typically exhibit more negative attitudes towards Catalan **(CITATIONS)**, which is consistent with our results.
+
+Lastly, province was also found as an important predictor, but only for some statements.
+In two of the solidarity statements (*És simpàtic* 'S/he is nice'; *És de fiar* 'S/he is trustworthy'), the speaker was rated more positively by those participants from the province of Barcelona.
+This finding might suggest that they identify with the speaker; thus, giving her higher ratings **BUSCAR CITA**.
+Interestingly, the speaker that created the stimuli is from the province of Girona, but has spent the last 11 years of her life living outside that area, either in Barcelona or in another country.
+We asked her to read the sentences using a Central Catalan 'neutral accent', which people might have potentially related with Barcelona.
+This 'neutral accent' does not show any of the prototypical characteristics of the *accent gironí*, a subvariety within Central Catalan spoken in some areas of Girona **(Gros, 2024)**.
+
+In summary, our findings suggest that the relationship between covert attitudes and speaker ratings is not straightforward, as the speaker was evaluated differently based on the social characteristic, with varying factors influencing such ratings (mother tongue and province).
+Crucially, phoneme production did not had an effect on the ratings, even though prescriptivist norms disfavor the realization of [s] in intervocalic position.
+Future studies should include a larger participant pool, as well as participants from other regions in Catalonia, such as Lleida and Tarragona.
+In addition, examining other phoneme productions that emerge through the contact of Catalan with Spanish could give us a better understanding of the factors that affect the attitudes towards the Catalan language, and potential sociolinguistic stereotypes and indicators related to Catalan identity.
 

### 2024-04-27 manuscript/main.tex

@@
 
-Write something here
+Catalan and Spanish are two languages that have coexisted for centuries,
+mainly in eastern Spain, along the Mediterranean cost \textbf{(Hualde,
+1992)}. It is safe to assume that there is a bidirectional influence
+between these languages: in that region, Catalan has influenced Spanish
+and Spanish has influenced Catalan in all linguistic levels
+\textbf{(Davidson, 2014, 2015)}. A crucial area of cross-linguistic
+influence might be the production of sounds, especially in those sounds
+who have different distribution in Catalan and Spanish. For example,
+previous studies have found that, in Barcelona, the production of
+Catalan vowels is different based on the L1 of the speaker, being it
+less target-like in L1 Spanish speakers compared to L1 Catalan speakers
+\textbf{(CITAR SEBASTIAN GALLES, etc)}.
+
+As for how Catalan people perceive these differences in production and,
+thus, their attitudes towards a less Catalan-like production of a given
+phoneme little is known. The aim of this study is to examine the covert
+attitudes towards the production of intervocalic /s/, as either {[}s{]}
+(disfavored by prescriptivist norms) or {[}z{]} (preferred by
+prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish
+bilinguals from Catalonia using a matched guise test \textbf{(Lambert,
+1967)}. The alveolar fricative was chosen because it exists in both
+Catalan and Spanish inventories, but their contextual distribution is
+different (Llisterri, 2023).
 
@@
 
-Both Catalan and Peninsular Spanish sound inventories have an phoneme
+Both Catalan and Peninsular Spanish sound inventories have a phoneme
 /s/, which can be classified as a voiceless alveolar fricative
@@
 phenomenon has mainly been attested in those areas where there has been
-an extended contact between Catalan and Spanish, such asthe so-called
+an extended contact between Catalan and Spanish, such as the so-called
 ``franja de Aragón'' (borderland between Catalonia and Aragon), some
@@
 speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
-origin' showed similar results, as illustrated in Figures \ref{fig:pbc}
-and \{fig:oc\}. Those who have both Catalan and Spanish as their L1
-rated the speaker more positively regardless of phoneme, compared to
-those whose L1 is either Catalan or Spanish.
+origin' show similar results, as illustrated in Figures \ref{fig:pbc}
+and \{fig:oc\}. Those who have both Catalan and Spanish as their L1 rate
+the speaker more positively regardless of phoneme, compared to those
+whose L1 is either Catalan or Spanish. In addition, although it is not
+statistically significant, participants give higher ratings to the
+speaker when they are listening to those statements with {[}z{]}.
 
@@
 both), and province of origin (Girona vs Barcelona) influenced the
-attitudes of the participants towards a Catalan native speaker. The
-results suggest that these predictors have a different effect on the
-ratings of the speaker depending on the social characteristic under
-evaluation.
+attitudes of the participants towards a native speaker of Central
+Catalan. The results suggest that these predictors have a different
+effect on the ratings of the speaker depending on the social
+characteristic under evaluation.
 
@@
 One explanation for this divergent pattern could be that either
-participants could not distinguish between {[}s{]} and {[}z{]} or, if
-they could, this difference in production was not changing their
-attitudes towards the speaker. This finding is interesting considering
-that, in our stimuli, the production of the voiceless alveolar fricative
-in intervocalic position has been considered a result of Catalan in
-contact with Spanish \textbf{(Benet et al., 2012; Davidson, 2020;
-Recasens Vives, 1996, p.~268)}, and, from a prescriptivist point of
-view, it should be avoided \textbf{(Institut d'Estudis Catalans, 2022)}.
-Despite the lack of significance of this predictor, it should be
-mentioned that the speaker received more positive ratings when the
-{[}z{]} was being produced. Although at this point we can only
-hypothesize, it could be that the combination of the phoneme /s/ with
-the precedent vowel, might give the listener more cues and have a
-stronger social meaning. Future studies should address this issue in
-order to explore if the combination of vowel + /s/ carries other social
-connotations. \textbf{REVISAR}
+participants could not distinguish between {[}s{]} and {[}z{]}
+\textbf{(Davidson, 2019)} or, if they could, this difference in
+production was not changing their attitudes towards the speaker
+\textbf{(Sinner, 2002)}. This finding is interesting considering that
+the production of the voiceless alveolar fricative in intervocalic
+position has been considered a result of Catalan in contact with Spanish
+\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
+p.~268)}, and, from a prescriptivist point of view, it should be avoided
+\textbf{(Institut d'Estudis Catalans, 2022)}. Despite the lack of
+significance of this predictor, it should be mentioned that the speaker
+received more positive ratings when the {[}z{]} was being produced. The
+fact that phoneme was not an important factor contrasts with
+\textbf{Davidson (2020)}, who analyzed covert and overt attitudes
+towards Catalonian Spanish and found that {[}z{]} was a significant
+predictor: it was considered a sociolinguistic indicator of
+Catalan-Spanish bilingual membership. Although at this point we can only
+hypothesize, it could be that {[}z{]} is perceptually more salient in
+Spanish than in Catalan, as the Spanish inventory does not have the
+phoneme /z/, but voicing of /s/ might partially occur \textbf{(Benet et
+al., 2012; Hualde \& Colina, 2015; Hualde et al., 2020)}. Another
+explanation could be that the /s/ alone does not provide enough social
+information; rather, the combination of the phoneme /s/ with the
+precedent vowel might give the listener more cues and have a stronger
+social meaning. Future studies should address this issue in order to
+explore if the combination of vowel + /s/ carries other social
+connotations than just the /s/ alone.
+
+We expected mother tongue to be a strong predictor of the attitudes
+towards the production of intervocalic /s/, specifically, that L1
+Catalan and bilingual speakers would give higher rates to those
+statements with {[}z{]}. This hypothesis was only partially confirmed.
+As reported above, phoneme was not a factor influencing the attitudes,
+and mother tongue was only a significant predictor in the two statements
+related to Catalan identity, namely \emph{Parla bé català} `S/he speaks
+proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
+origin'. Interestingly, bilinguals gave higher ratings to the speaker,
+followed by L1 Catalan participants and L1 Spanish participants. A
+potential explanation for this finding, that should be verified in
+future research, could be that L1 Catalan speakers have specific ideas
+(higher standards) towards the `proper' Catalan. In addition, some
+studies have suggested that those Catalan people who grow up in
+Spanish-speaking households typically exhibit more negative attitudes
+towards Catalan \textbf{(CITATIONS)}, which is consistent with our
+results.
+
+Lastly, province was also found as an important predictor, but only for
+some statements. In two of the solidarity statements (\emph{És simpàtic}
+`S/he is nice'; \emph{És de fiar} `S/he is trustworthy'), the speaker
+was rated more positively by those participants from the province of
+Barcelona. This finding might suggest that they identify with the
+speaker; thus, giving her higher ratings \textbf{BUSCAR CITA}.
+Interestingly, the speaker that created the stimuli is from the province
+of Girona, but has spent the last 11 years of her life living outside
+that area, either in Barcelona or in another country. We asked her to
+read the sentences using a Central Catalan `neutral accent', which
+people might have potentially related with Barcelona. This `neutral
+accent' does not show any of the prototypical characteristics of the
+\emph{accent gironí}, a subvariety within Central Catalan spoken in some
+areas of Girona \textbf{(Gros, 2024)}.
+
+In summary, our findings suggest that the relationship between covert
+attitudes and speaker ratings is not straightforward, as the speaker was
+evaluated differently based on the social characteristic, with varying
+factors influencing such ratings (mother tongue and province).
+Crucially, phoneme production did not had an effect on the ratings, even
+though prescriptivist norms disfavor the realization of {[}s{]} in
+intervocalic position. Future studies should include a larger
+participant pool, as well as participants from other regions in
+Catalonia, such as Lleida and Tarragona. In addition, examining other
+phoneme productions that emerge through the contact of Catalan with
+Spanish could give us a better understanding of the factors that affect
+the attitudes towards the Catalan language, and potential
+sociolinguistic stereotypes and indicators related to Catalan identity.
 

### 2024-04-29 manuscript/main.Rmd
update

@@
 
-Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean cost **(Hualde, 1992)**.
+Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean coast **(Hualde, 1992)**.
 It is safe to assume that there is a bidirectional influence between these languages: in that region, Catalan has influenced Spanish and Spanish has influenced Catalan in all linguistic levels **(Davidson, 2014, 2015)**.
-A crucial area of cross-linguistic influence might be the production of sounds, especially in those sounds who have different distribution in Catalan and Spanish.
+A crucial area of cross-linguistic influence might be the production of sounds, especially in those sounds that have different distribution in Catalan and Spanish.
 For example, previous studies have found that, in Barcelona, the production of Catalan vowels is different based on the L1 of the speaker, being it less target-like in L1 Spanish speakers compared to L1 Catalan speakers **(CITAR SEBASTIAN GALLES, etc)**.
 
-As for how Catalan people perceive these differences in production and, thus, their attitudes towards a less Catalan-like production of a given phoneme little is known.
-The aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test **(Lambert, 1967)**.
+However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound.
+Therefore, the aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test **(Lambert, 1967)**.
 The alveolar fricative was chosen because it exists in both Catalan and Spanish inventories, but their contextual distribution is different **(Llisterri, 2023)**.
-
-**PONER ALGO MÁS**
 
@@
 
-Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process \cite{hualde_colina_2015}.
-Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) **(Hualde & Colina, 2015; Hualde et al., 2020)**.
-Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process.
+\cite{hualde_colina_2015
+
+While Spanish inventory does not have the phoneme /z/, the sound [z] may occur as an allophone of /s/, due to a voice assimilation process.
 It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.
-
 There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead.
@@
 
-In summary, both Catalan and Spanish have a voiceless alveolar fricative in their inventories.
-Catalan also has a voiced variant, which is produced when this sound is followed by a voiced consonant or a vowel.
-Spanish might produced this sound only when it is followed by a voiced consonant, although /s/ might be "partially voiced" (Hualde & Colina, 2015).
-Thus, these phonemes, although they have the same allophones, their contextual distribution is different (Llisterri, 2023).
-Moreover, some studies have shown that there is a bidirectional influence between both languages in the production of the voiced vs the voiceless alveolar fricative, and that such influence might be different based on the context.
-These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.
-
 \subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}
@@
 The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual **(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003)** and bilingual contexts **(Davidson, 2014, 2015, 2019; Sinner, 2002)**, despite having been categorized as a low-salient variant **(Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)**.
-With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
+With respect to the former, it has been found that for listeners from Costa Rica and Ecuador (García, 2019), the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
 Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine.
 Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class **(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003)**.
+Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects that exhibit systematic voicing.
 
-Similarly, **García (2019)** shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers.
-Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older.
-Contrary, men who produce intervocalic [z] do not receive these negative evaluations.
-Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing.
-**García (2020)** hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties.
-Interestingly, the mean voicing rates of females are roughly two age groups behind those of males.
-Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.
-
-With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan **(Davidson, 2014, 2015)**.
+With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations **(Davidson, 2014, 2015)**.
 According to **Sinner (2002)**, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness **(Davidson, 2019; Sinner, 2002)**.
-In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish.
-Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not.
 Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization.
@@
 \subsection{The present study}
-**INTRO**.
 
-Accordingly, the research questions and hypotheses are as follows:
+Accordingly, and considering that there are no perception studies focusing on Catalonian Catalan (central variety), the present study aims to answer the following research questions:
+
 1. What covert attitudes are linked to the perception Catalan intervocalic /s/ as either [s] or [z]?
-2. What is the relationship between  individual social factors (mother tongue, province of origin) and covert attitudes towards Catalan intervocalic /s/?
+
+2. What is the relationship between  individual social factors and covert attitudes towards Catalan intervocalic /s/?
 
@@
 The first part, a demographic questionnaire, collected information regarding participant age, gender, place of origin, place of residence, mother tongue, family income, and level of studies.
-The second portion of the survey consisted of a perception task where participants were presented with a series of stimuli containing the target phoneme contrast (intervocalic voiced and voiceless alveolar fricative) followed by 11 evaluative statements related to social characteristics such as level of studies, friendliness, social class, and ethnic origin.
+The second portion of the survey consisted of a perception task where participants were presented with a series of stimuli containing the target phoneme contrast (intervocalic voiced and voiceless alveolar fricative) such as *han tancat les vies estretes* 'they closed the narrow streets', followed by 11 evaluative statements related to social characteristics such as level of studies, friendliness, social class, and ethnic origin.
 These evaluative statements were then followed by a 100-point draggable sliding scale with which participants indicated their level of agreement or disagreement with the content of the statement, where 0 was “totally disagree” and 100 was “totally agree”.
@@
 Participants self-identified for this study.
-Recruitment materials in Catalan with information about the study, the requirements to participate and a link to the survey were posted on different social networks and online communication channels, such as WhatsApp and Instagram.
-Participants who were interested and wanted to participate clicked on the link and read the study description.
-Those who decided to participate completed the survey on their own electronic devices and at their own pace without being observed.
-On average, the survey took approximately 15 minutes to be completed.
-The survey followed a between-subjects design.
+These were the requirements: to be over 18 years of age, to be from Catalonia, and to be proficient in both Catalan and Spanish.
+The survey was shared using social networks and online communication channels and it followed a between-subjects design.
 Therefore, participants were randomly assigned to one of the two versions.
@@
 Data are drawn from a total of 22 participants (13 females) aged between 18 and 72 years old (mean = 38) from Barcelona and Girona.
-In order to be eligible to take part in the study, participants had to meet the following three requirements: to be over 18 years of age, to be from Catalonia, and to be proficient in both Catalan and Spanish.
 
@@
 Data were analyzed in R (R Core Team, 2023).
-First, we performed a descriptive analysis that consisted of creating a series of plots and tables to summarize and describe the features of the dataset and to identify potential trends.
-After that, we performed inferential statistical analysis.
 For each of the 11 evaluative statements, we created a series of hierarchically nested linear regression models using the lme4 package (Bates et al., 2015) with attitudinal sliding scale responses as the dependent variable, and phoneme, mother tongue, and province as potential predictors.
@@
 
-Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' show similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}.
+Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' show similar results, as illustrated in Figure \ref{fig:oc}.
 Those who have both Catalan and Spanish as their L1 rate the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish.
 In addition, although it is not statistically significant, participants give higher ratings to the speaker when they are listening to those statements with [z].
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=15cm]{./includes/figures/pbc.png}
-\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
-\end{center}
-\end{figure}
 
@@
 
-Participant responses to the statements *És simpàtic* 'S/he is nice' and *És de fiar* 'S/he is trustworthy' showed similar results, as illustrated in Figures \ref{fig:simpatic} and {fig:fiar}.
+Participant responses to the statements *És simpàtic* 'S/he is nice' and *És de fiar* 'S/he is trustworthy' showed similar results, as illustrated in Figures \ref{fig:fiar}.
 Those who are from and live in the province of Barcelona give higher ratings to the speaker, compared to those who are from and live in Girona.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=15cm]{./includes/figures/simpatic.png}
-\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
-\end{center}
-\end{figure}
 
@@
 
-\section{Discussion and Conclusions}
-The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a native speaker of Central Catalan.
+\section{Discussion}
+
+The aim of this exploration study was to explore the covert attitudes that are linked to the perception Catalan intervocalic /s/ as either [s] or [z] and to examine the relationship between  individual social factors and covert attitudes towards Catalan intervocalic /s/.
 The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.
@@
 
+\section{Conclusion}
+
 In summary, our findings suggest that the relationship between covert attitudes and speaker ratings is not straightforward, as the speaker was evaluated differently based on the social characteristic, with varying factors influencing such ratings (mother tongue and province).
@@
 
-\subsubsection{Headings}
-
-Section headings are centered in boldface with capitalized letters.
-Sub-headings start at the left margin in the column with the first letter capitalized and the rest of the heading in lower case.
-Sub-sub-headings appear like sub-headings, except that they are in italics and not boldface.
-See examples in this file.
-No more than 3 levels of headings should be used.
-Empty lines should be left above and below each section heading.
-
-\subsection{Text font}
-
-Times or Times New Roman font is used for the main text.
-Recommended font size is 11 points.
-Other font types may be used if needed for special purposes.
-If using any non-Unicode fonts, these must be embedded in the final PDF file.
-
-The \LaTeX{} template can be used either with plain \LaTeX{} or \XeLaTeX.
-
-\subsection{Figures}
-
-All figures should be centered on the column (or page, if the figure
-spans both columns). Figure captions should follow each figure and
-have the format given in Fig. \ref{fig:vowels}.
-
-Figures should preferably be line drawings. If they contain grey
-shades or colours, it should be checked that they print well on a
-high-quality noncolour laser printer.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6cm]{./includes/figures/ipa.eps}
-\caption{The vowel chart used in the International Phonetic
-Alphabet (IPA).}\label{fig:vowels}
-\end{center}
-\end{figure}
-
-\subsection{Tables}
-
-An example of a table is shown as Table~\ref{tab:decibel}.  Somewhat
-different styles are allowed according to the type and purpose of the
-table. Colour should not be used, but grey shading is allowed. There
-should be a margin of 6~points (pt) above and below the table.
-
-The caption text may be above or below the table, but this should be
-consistent throughout the submission. Left and right indentation of
-the caption should be 0.5~cm.
-
-\begin{table}[!ht]
-  \begin{center}
-  \begin{tabular}{|c|c|}
-  \hline
-  \rowcolor[gray]{.75}
-  ratio    & Decibels\\
-  \hline
-  $1/1$    & $0$\\
-  $2/1$    & $6$\\
-  $3.16$   & $10$\\
-  $1/10$   & $-20$\\
-  $10/1$   & $20$\\
-  $100/1$  & $40$\\
-  $1000/1$ & $60$\\
-  \hline
-  \end{tabular}
-  \caption{This is an example of a table showing Decibel (dB)
-  ratios.}\label{tab:decibel}
-  \end{center}
-\end{table}
-
-\subsection{Equations}
-
-Equations should be placed on separate lines and numbered. An example
-of an equation is
-\begin{equation}\label{eq:tzero}
-T_0 = \frac{1}{f_0}.
-\end{equation}
-Numbers of equations can be on the right or on the left margin of the
-text column.
-
-\subsection{Examples}
-
-Examples from other languages can either be presented in the body text, or, if referred to elsewhere or particularly long and complex, can be put on a separate, numbered line, as should be done for equations.
-
-\subsection{Phonetic fonts}
-
-We recommend that you use the \textsc{tipa} package for IPA phonetic symbols.
-For information about how to access Unicode fonts in the Word template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}.
-The font you use must be embedded.
-Please remember to check this,  e.g.
-by inspecting the ``Document Properties --- Fonts'' in Acrobat Reader.
-
-\subsection{Page numbering}
-
-Page numbers will be added electronically to the document
-later. \textit{Please do not add page numbers and please do not insert
-any footers or headers!}
-
-\subsection{References}
-
-Please use just the reference number in square brackets. Formulations
-with author names like "\ldots\ as Ladefoged \cite{Ladefoged:2003}
-showed that \ldots" are acceptable but not "as shown in [Ladefoged,
-3]" or "as shown in (Ladefoged [3])".
-
-References are to be numbered in alphabetical order. Please
-double-check the final version of your paper with regard to the
-correct correspondence of references to their numbers.
-
-\subsection{Hyperlinks}
-
-Links to URLs or email addresses should be formatted as normal text,
-\textit{not} as hyperlinks and not blue or underlined etc. Usually
-hyperlinks to web pages are listed in the references section. If
-required, line breaks can be placed within URLs after slashes or
-dashes (cf.~\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that no hyphens
-are inserted.
-
-\subsection{Footnotes and endnotes}
-
-If footnotes cannot be avoided they should appear as
-endnotes.\footnote{This footnote appears here as an endnote.}
-
-\section{Multimedia files}
-
-Multimedia data that are part of the paper are to be embedded in the submitted PDF; they cannot be submitted as supplementary data.
-Any images are to be included in the paper as Figures (see Section 2.3 above).
-It is the authors' responsibility to check image quality ahead of submission.
-Audio examples are to be embedded within the PDF.
-To do this, authors can generate the PDF, and then embed the audio files using Adobe Acrobat Professional, or other software that offers the same outcome, so that the audio is included in the PDF.
-The presence of audio data should be identified in the text.
-
-We encourage authors to illustrate video data using still photographs from the video, and to include them as figures in the PDF.
-We cannot accept embedded video files, but authors are welcome to refer readers to a URL on the internet where these can be viewed.
-
-\section{PDF details}
-
-PDF files submitted must comply with the following requirements:
-
-\begin{enumerate}
-\item all special fonts and symbols must be embedded in the PDF
-  file so that correct rendering of the PDF does not depend on the
-  fonts installed on the viewer's computer
-\item there must be no
-  password protection on the PDF file, i.e. PDF files must not be
-  protected by PDF security in any way, i.e. content extraction,
-  document assembly, high resolution printing etc. must not be
-  forbidden
-\item PDF files should not contain any colours, hyperlinks,
-  multimedia or 3D content, and no JavaScript or forms
-\item PDF files
-  should be no larger than 5 MB.
-\end{enumerate}
-
-\section{Anonymity}
-
-In ICPhS 2023 submissions, an anonymous reviewing process will be used. This means that for the first submission the name(s) of the author(s) and
-their affiliation(s) \emph{must not} be mentioned.
-In addition, please refrain from using acknowledgements.
-Please also try to make your own previous research as anonymous as possible.
-As an example: do not write "In our previous study \cite{Stevens:1999} we could show\ldots" but "As shown in \cite{Stevens:1999}\ldots".
-Or refer to your own published or otherwise widely known work, and to that of the other authors, in the "Julius Caesar style", i.e.
-in the third person (for example: his work, her work, their work).
-Reference as "anonymous" only work that you or the other authors have submitted for publication, but which has not yet been published, e.g.
-\cite{Anonymous:submitted}.
-
-Please make sure that no author details appear in the Document Properties of the PDF file.
-\emph{For the revised paper submission author details are of course needed}.
-Acknowledgements and references to one's own work are possible as usual.
-
-\section{Format of references}
-
-References are formatted using the IEEE standard (available in various reference management systems like Zotero or Mendeley).
-If you do not use a reference management system, please use the examples provided for monographs \cite{Fant:1960}, contributions to volumes \cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles in conference proceedings \cite{Ladefoged:2003}.
-Abbreviations of
-well-known conferences and journals are possible \cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016} are referenced according to authors' instructions.
-
-\section{R scripts}
-
-You can use knitr code chunks like in any .Rmd document:
-
-```{r, 'chunk_ex'}
-2 + 2
-```
-
-This includes inline code: 2 + 2 = `r 2 + 2`.
-
-R scripts can be loaded as well:
-
-```{r, 'load_script', echo=TRUE, message=FALSE, warning=FALSE}
-source("./includes/scripts/analysis.R")
-```
-
-Which means you have access to any objects assigned in the script, like Figure 
-\ref{fig:ex_plot}:
-
-```{r, 'ex_plot', fig.align='center', fig.width=2.75, fig.height=1.5, fig.cap.top="This is a figure caption.", fig.pos='h', fig.lp='fig:'} 
-cars_plot
-```
-
 \bibliographystyle{./includes/bib/IEEEtran.bst}
 \bibliography{./includes/bib/icphs2023.bib}
-
-\theendnotes

### 2024-04-29 manuscript/includes/tex/doc_prefix.tex

\title{Language Attitudes of Catalan-Spanish Bilinguals Towards Intervocalic Alveolar Fricatives in Catalan}
\author{Please write XXXX instead of the name(s) of the author(s)}
\organization{Please write XXXX instead of the affiliation(s)}
\email{please write XXXX instead of the email address(es)}

\maketitle

\begin{abstract}
Catalan and Spanish are two languages that have coexisted for centuries in Eastern Spain. Thus, there is a bidirectional influence between these languages in all linguistic levels, including in the production of sounds. However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound. Therefore, the aim of this study is to examine the covert attitudes toward the production of intervocalic /s/, as either [s] or [z] by Catalan-Spanish bilinguals from Catalonia using a matched guise test. Results indicate that phoneme type is not a significant predictor of language attitudes in Catalonia, but other individual social factors, such as mother tongue and province, might be.
\end{abstract}

\keywords{Catalan, Spanish, language attitudes, intervocalic alveolar fricatives.}

### 2024-04-29 manuscript/main.tex

@@
 \usepackage{xcolor}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{48,48,48}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{\textbf{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.87,0.87,0.75}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.86,0.86,0.80}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.76,0.75,0.62}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.75,0.75,0.82}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.94,0.94,0.82}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{\textbf{#1}}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
 \usepackage{graphicx}
@@
 
-\title{Paper template for {ICPhS} 2023 Prague}
-\author{Please write XXX instead of the name(s) of the author(s)}
-\organization{Please write XXX instead of the affiliation(s)}
-\email{please write XXX instead of the email address(es)}
+\title{Language Attitudes of Catalan-Spanish Bilinguals Towards Intervocalic Alveolar Fricatives in Catalan}
+\author{Please write XXXX instead of the name(s) of the author(s)}
+\organization{Please write XXXX instead of the affiliation(s)}
+\email{please write XXXX instead of the email address(es)}
 
@@
 \begin{abstract}
-This is the layout specification and template definition for the papers of ICPhS XX (the 20th International Congress of Phonetic Sciences, which will be held in Prague, Czech Republic, August 7-11, 2023).
-This template is a revised version of the one used for Melbourne ICPhS XIX in 2019, originally generated from the template for Speech Prosody 2006 in Dresden.
-
-The abstract may consist of more than one paragraph but must be kept within a 150-word limit. 
-This abstract will be printed in the abstract booklet to be given out at the conference.
+Catalan and Spanish are two languages that have coexisted for centuries in Eastern Spain.
+Thus, there is a bidirectional influence between these languages in all linguistic levels, including in the production of sounds.
+However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound.
+Therefore, the aim of this study is to examine the covert attitudes toward the production of intervocalic /s/, as either [s] or [z] by Catalan-Spanish bilinguals from Catalonia using a matched guise test.
+Results indicate that phoneme type is not a significant predictor of language attitudes in Catalonia, but other individual social factors, such as mother tongue and province, might be.
 \end{abstract}
 
-\keywords{There is space for up to five self-selected keywords (maximally two 
-lines).}
+\keywords{Catalan, Spanish, language attitudes, intervocalic alveolar fricatives.}
 
@@
 Catalan and Spanish are two languages that have coexisted for centuries,
-mainly in eastern Spain, along the Mediterranean cost \textbf{(Hualde,
+mainly in eastern Spain, along the Mediterranean coast \textbf{(Hualde,
 1992)}. It is safe to assume that there is a bidirectional influence
@@
 influence might be the production of sounds, especially in those sounds
-who have different distribution in Catalan and Spanish. For example,
+that have different distribution in Catalan and Spanish. For example,
 previous studies have found that, in Barcelona, the production of
@@
 
-As for how Catalan people perceive these differences in production and,
-thus, their attitudes towards a less Catalan-like production of a given
-phoneme little is known. The aim of this study is to examine the covert
-attitudes towards the production of intervocalic /s/, as either {[}s{]}
-(disfavored by prescriptivist norms) or {[}z{]} (preferred by
-prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish
-bilinguals from Catalonia using a matched guise test \textbf{(Lambert,
-1967)}. The alveolar fricative was chosen because it exists in both
-Catalan and Spanish inventories, but their contextual distribution is
-different (Llisterri, 2023).
+However, little is known about the perception and the attitudes toward a
+less Catalan-like production of a given sound. Therefore, the aim of
+this study is to examine the covert attitudes towards the production of
+intervocalic /s/, as either {[}s{]} (disfavored by prescriptivist norms)
+or {[}z{]} (preferred by prescriptivist norms) by L1 Catalan, L1
+Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched
+guise test \textbf{(Lambert, 1967)}. The alveolar fricative was chosen
+because it exists in both Catalan and Spanish inventories, but their
+contextual distribution is different \textbf{(Llisterri, 2023)}.
 
@@
 
-Interestingly, Medieval Spanish, similarly to Modern Catalan,
-distinguished between /s/ and /z/ in intervocalic position within a
-word, although these sounds merged through a devoicing process
-\cite{hualde_colina_2015}. Nevertheless, in those modern varieties where
-the /s/ in coda position is produced, such as in Peninsular Spanish,
-this voiceless phoneme might assimilate to the following sound,
-especially if it is a voiced consonant (\emph{mismo}, {[}mízmo{]}),
-rather than a vowel (\emph{las alas}, {[}lasálas{]}) \textbf{(Hualde \&
-Colina, 2015; Hualde et al., 2020)}. Hence, Spanish inventory does not
-have the phoneme /z/, but {[}z{]} may occur as an allophone of /s/, due
-to a voice assimilation process. It is important to mention, though,
-that it has been claimed that this voicing is generally not complete,
-but partial \textbf{(Benet et al., 2012; Hualde \& Colina, 2015; Hualde
-et al., 2020)}.
+\textbackslash cite\{hualde\_colina\_2015
 
-There is some evidence of the influence of Spanish on the production of
-the Catalan alveolar fricative, in contexts where, according to
-prescriptivist rules \textbf{(Institut d'Estudis Catalans, 2022)}, the
-{[}z{]} should be produced, but {[}s{]} is produced instead. This
-phenomenon has mainly been attested in those areas where there has been
-an extended contact between Catalan and Spanish, such as the so-called
-``franja de Aragón'' (borderland between Catalonia and Aragon), some
-areas of the Valencian Community, and Barcelona and other urban areas
-\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
-p.~268)}. The opposite pattern has also been observed - that is, the
-voicing of the alveolar fricative in the Spanish spoken in Barcelona and
-Valencia. Interestingly, although in both Barcelona and Valencia Catalan
-and Spanish coexist and influence each other, it seems that Catalan has
-a greater influence on Spanish in Barcelona, but Spanish has a greater
+While Spanish inventory does not have the phoneme /z/, the sound {[}z{]}
+may occur as an allophone of /s/, due to a voice assimilation process.
+It is important to mention, though, that it has been claimed that this
+voicing is generally not complete, but partial \textbf{(Benet et al.,
+2012; Hualde \& Colina, 2015; Hualde et al., 2020)}. There is some
+evidence of the influence of Spanish on the production of the Catalan
+alveolar fricative, in contexts where, according to prescriptivist rules
+\textbf{(Institut d'Estudis Catalans, 2022)}, the {[}z{]} should be
+produced, but {[}s{]} is produced instead. This phenomenon has mainly
+been attested in those areas where there has been an extended contact
+between Catalan and Spanish, such as the so-called ``franja de Aragón''
+(borderland between Catalonia and Aragon), some areas of the Valencian
+Community, and Barcelona and other urban areas \textbf{(Benet et al.,
+2012; Davidson, 2020; Recasens Vives, 1996, p.~268)}. The opposite
+pattern has also been observed - that is, the voicing of the alveolar
+fricative in the Spanish spoken in Barcelona and Valencia.
+Interestingly, although in both Barcelona and Valencia Catalan and
+Spanish coexist and influence each other, it seems that Catalan has a
+greater influence on Spanish in Barcelona, but Spanish has a greater
 influence on Catalan in Valencia \textbf{(Davidson, 2020)}.
-
-In summary, both Catalan and Spanish have a voiceless alveolar fricative
-in their inventories. Catalan also has a voiced variant, which is
-produced when this sound is followed by a voiced consonant or a vowel.
-Spanish might produced this sound only when it is followed by a voiced
-consonant, although /s/ might be ``partially voiced'' (Hualde \& Colina,
-2015). Thus, these phonemes, although they have the same allophones,
-their contextual distribution is different (Llisterri, 2023). Moreover,
-some studies have shown that there is a bidirectional influence between
-both languages in the production of the voiced vs the voiceless alveolar
-fricative, and that such influence might be different based on the
-context. These divergent patterns might be explained by the different
-sociopolitic and linguistic realities of each city, and might have an
-effect on speakers' attitudes towards the devoicing of the alveolar
-fricative /z/ in Catalan.
 
@@
 2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
-former, it has been found that for listeners from San José, Costa Rica,
-the covert attitudes associated with intervocalic {[}z{]} vary
-significantly by speaker gender. Consequently, male speakers who use
-{[}z{]} tend to be positively evaluated, and generally perceived as
-nice, local, confident, and masculine. Conversely, female speakers who
-use {[}z{]} are only afforded negative associations of lower education
-and social class \textbf{(Chappell, 2016; Chappell et al., 2023;
-Silverstein, 2003)}.
-
-Similarly, \textbf{García (2019)} shows that, in Ecuador, intervocalic
-/s/ is evaluated differently in male and female speakers. Accordingly,
-women who produce intervocalic {[}z{]} are viewed as significantly lower
-status, less pleasant, and older. Contrary, men who produce intervocalic
-{[}z{]} do not receive these negative evaluations. Additionally, this
-variable has been found to serve as a marker of regional identity, with
-voicing being associated with neighboring Highland Ecuadorian dialects
-(Cuenca and Quito) that exhibit systematic voicing. \textbf{García
-(2020)} hypothesizes that young men are the leaders of this change,
-essentially due to the fact that they are in greater contact with these
-highland varieties. Interestingly, the mean voicing rates of females are
-roughly two age groups behind those of males. Consequently, it is
-believed that it is not necessarily the case that women in Ecuador are
-avoiding voicing like San José women due to its negative evaluations,
-but rather that they are further behind in the adoption of the incoming
-variant.
+former, it has been found that for listeners from Costa Rica and Ecuador
+(García, 2019), the covert attitudes associated with intervocalic
+{[}z{]} vary significantly by speaker gender. Consequently, male
+speakers who use {[}z{]} tend to be positively evaluated, and generally
+perceived as nice, local, confident, and masculine. Conversely, female
+speakers who use {[}z{]} are only afforded negative associations of
+lower education and social class \textbf{(Chappell, 2016; Chappell et
+al., 2023; Silverstein, 2003)}. Additionally, this variable has been
+found to serve as a marker of regional identity, with voicing being
+associated with neighboring Highland Ecuadorian dialects that exhibit
+systematic voicing.
 
@@
 {[}z{]} in bilingual contexts, this phenomenon has been mainly studied
-in Catalan-Spanish contact situations, given its robust usage by
-Catalan-Spanish bilinguals and, in particular, by those with greatest
-exposure to and usage of Catalan \textbf{(Davidson, 2014, 2015)}.
+in Catalan-Spanish contact situations \textbf{(Davidson, 2014, 2015)}.
 According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
@@
 its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
-2002)}. In their study, Sinner (2002) interviewed 12 speakers of
-Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27
-to 41 regarding their awareness of linguistic features of Catalanized
-Spanish. Results showed that only select Barcelona speakers perceived
-intervocalic {[}z{]} as a linguistic marker of Catalonian Spanish,
-whereas monolingual Spanish-speakers from Madrid did not. Additionally,
-Davidson (2019) found that, in Barcelona, the indexical field of
-intervocalic {[}z{]} features exclusively positive associations with
-bilingualism and local solidarity, which warrants its gradual adoption
-in the speech community, unlike other phonetic markers of Catalonian
-Spanish that are the topic of overt social ridicule, such as lateral
-velarization. These results are in line with previous findings on
-intervocalic {[}z{]} in Catalonian Spanish. For instance,
+2002)}. Additionally, Davidson (2019) found that, in Barcelona, the
+indexical field of intervocalic {[}z{]} features exclusively positive
+associations with bilingualism and local solidarity, which warrants its
+gradual adoption in the speech community, unlike other phonetic markers
+of Catalonian Spanish that are the topic of overt social ridicule, such
+as lateral velarization. These results are in line with previous
+findings on intervocalic {[}z{]} in Catalonian Spanish. For instance,
 \textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
@@
 
-\textbf{INTRO}. Accordingly, the research questions and hypotheses are
-as follows: 1. What covert attitudes are linked to the perception
-Catalan intervocalic /s/ as either {[}s{]} or {[}z{]}? 2. What is the
-relationship between individual social factors (mother tongue, province
-of origin) and covert attitudes towards Catalan intervocalic /s/?
+Accordingly, and considering that there are no perception studies
+focusing on Catalonian Catalan (central variety), the present study aims
+to answer the following research questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  What covert attitudes are linked to the perception Catalan
+  intervocalic /s/ as either {[}s{]} or {[}z{]}?
+\item
+  What is the relationship between individual social factors and covert
+  attitudes towards Catalan intervocalic /s/?
+\end{enumerate}
 
@@
 the target phoneme contrast (intervocalic voiced and voiceless alveolar
-fricative) followed by 11 evaluative statements related to social
+fricative) such as \emph{han tancat les vies estretes} `they closed the
+narrow streets', followed by 11 evaluative statements related to social
 characteristics such as level of studies, friendliness, social class,
@@
 
-Participants self-identified for this study. Recruitment materials in
-Catalan with information about the study, the requirements to
-participate and a link to the survey were posted on different social
-networks and online communication channels, such as WhatsApp and
-Instagram. Participants who were interested and wanted to participate
-clicked on the link and read the study description. Those who decided to
-participate completed the survey on their own electronic devices and at
-their own pace without being observed. On average, the survey took
-approximately 15 minutes to be completed. The survey followed a
+Participants self-identified for this study. These were the
+requirements: to be over 18 years of age, to be from Catalonia, and to
+be proficient in both Catalan and Spanish. The survey was shared using
+social networks and online communication channels and it followed a
 between-subjects design. Therefore, participants were randomly assigned
@@
 drawn from a total of 22 participants (13 females) aged between 18 and
-72 years old (mean = 38) from Barcelona and Girona. In order to be
-eligible to take part in the study, participants had to meet the
-following three requirements: to be over 18 years of age, to be from
-Catalonia, and to be proficient in both Catalan and Spanish.
+72 years old (mean = 38) from Barcelona and Girona.
 
@@
 
-Data were analyzed in R (R Core Team, 2023). First, we performed a
-descriptive analysis that consisted of creating a series of plots and
-tables to summarize and describe the features of the dataset and to
-identify potential trends. After that, we performed inferential
-statistical analysis. For each of the 11 evaluative statements, we
-created a series of hierarchically nested linear regression models using
-the lme4 package (Bates et al., 2015) with attitudinal sliding scale
-responses as the dependent variable, and phoneme, mother tongue, and
-province as potential predictors. To determine the best fit model for
-each statement, we used the anova() function (Fox \& Weisberg, 2019).
+Data were analyzed in R (R Core Team, 2023). For each of the 11
+evaluative statements, we created a series of hierarchically nested
+linear regression models using the lme4 package (Bates et al., 2015)
+with attitudinal sliding scale responses as the dependent variable, and
+phoneme, mother tongue, and province as potential predictors. To
+determine the best fit model for each statement, we used the anova()
+function (Fox \& Weisberg, 2019).
 
@@
 speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
-origin' show similar results, as illustrated in Figures \ref{fig:pbc}
-and \{fig:oc\}. Those who have both Catalan and Spanish as their L1 rate
-the speaker more positively regardless of phoneme, compared to those
-whose L1 is either Catalan or Spanish. In addition, although it is not
-statistically significant, participants give higher ratings to the
-speaker when they are listening to those statements with {[}z{]}.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=15cm]{./includes/figures/pbc.png}
-\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
-\end{center}
-\end{figure}
+origin' show similar results, as illustrated in Figure \ref{fig:oc}.
+Those who have both Catalan and Spanish as their L1 rate the speaker
+more positively regardless of phoneme, compared to those whose L1 is
+either Catalan or Spanish. In addition, although it is not statistically
+significant, participants give higher ratings to the speaker when they
+are listening to those statements with {[}z{]}.
 
@@
 nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
-results, as illustrated in Figures \ref{fig:simpatic} and \{fig:fiar\}.
-Those who are from and live in the province of Barcelona give higher
-ratings to the speaker, compared to those who are from and live in
-Girona.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[height=15cm]{./includes/figures/simpatic.png}
-\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
-\end{center}
-\end{figure}
+results, as illustrated in Figures \ref{fig:fiar}. Those who are from
+and live in the province of Barcelona give higher ratings to the
+speaker, compared to those who are from and live in Girona.
 
@@
 
-\section{Discussion and Conclusions}
+\section{Discussion}
 
-The aim of this exploration study was to investigate whether phoneme
-production ({[}s{]} vs {[}z{]}), mother tongue (Catalan vs Spanish vs
-both), and province of origin (Girona vs Barcelona) influenced the
-attitudes of the participants towards a native speaker of Central
-Catalan. The results suggest that these predictors have a different
-effect on the ratings of the speaker depending on the social
-characteristic under evaluation.
+The aim of this exploration study was to explore the covert attitudes
+that are linked to the perception Catalan intervocalic /s/ as either
+{[}s{]} or {[}z{]} and to examine the relationship between individual
+social factors and covert attitudes towards Catalan intervocalic /s/.
+The results suggest that these predictors have a different effect on the
+ratings of the speaker depending on the social characteristic under
+evaluation.
 
@@
 
+\section{Conclusion}
+
 In summary, our findings suggest that the relationship between covert
@@
 
-\subsubsection{Headings}
-
-Section headings are centered in boldface with capitalized letters.
-Sub-headings start at the left margin in the column with the first
-letter capitalized and the rest of the heading in lower case.
-Sub-sub-headings appear like sub-headings, except that they are in
-italics and not boldface. See examples in this file. No more than 3
-levels of headings should be used. Empty lines should be left above and
-below each section heading.
-
-\subsection{Text font}
-
-Times or Times New Roman font is used for the main text. Recommended
-font size is 11 points. Other font types may be used if needed for
-special purposes. If using any non-Unicode fonts, these must be embedded
-in the final PDF file.
-
-The \LaTeX{} template can be used either with plain \LaTeX{} or
-\XeLaTeX.
-
-\subsection{Figures}
-
-All figures should be centered on the column (or page, if the figure
-spans both columns). Figure captions should follow each figure and have
-the format given in Fig. \ref{fig:vowels}.
-
-Figures should preferably be line drawings. If they contain grey shades
-or colours, it should be checked that they print well on a high-quality
-noncolour laser printer.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6cm]{./includes/figures/ipa.eps}
-\caption{The vowel chart used in the International Phonetic
-Alphabet (IPA).}\label{fig:vowels}
-\end{center}
-\end{figure}
-
-\subsection{Tables}
-
-An example of a table is shown as
-Table\textasciitilde{}\ref{tab:decibel}. Somewhat different styles are
-allowed according to the type and purpose of the table. Colour should
-not be used, but grey shading is allowed. There should be a margin of
-6\textasciitilde points (pt) above and below the table.
-
-The caption text may be above or below the table, but this should be
-consistent throughout the submission. Left and right indentation of the
-caption should be 0.5\textasciitilde cm.
-
-\begin{table}[!ht]
-  \begin{center}
-  \begin{tabular}{|c|c|}
-  \hline
-  \rowcolor[gray]{.75}
-  ratio    & Decibels\\
-  \hline
-  $1/1$    & $0$\\
-  $2/1$    & $6$\\
-  $3.16$   & $10$\\
-  $1/10$   & $-20$\\
-  $10/1$   & $20$\\
-  $100/1$  & $40$\\
-  $1000/1$ & $60$\\
-  \hline
-  \end{tabular}
-  \caption{This is an example of a table showing Decibel (dB)
-  ratios.}\label{tab:decibel}
-  \end{center}
-\end{table}
-
-\subsection{Equations}
-
-Equations should be placed on separate lines and numbered. An example of
-an equation is \begin{equation}\label{eq:tzero}
-T_0 = \frac{1}{f_0}.
-\end{equation} Numbers of equations can be on the right or on the left
-margin of the text column.
-
-\subsection{Examples}
-
-Examples from other languages can either be presented in the body text,
-or, if referred to elsewhere or particularly long and complex, can be
-put on a separate, numbered line, as should be done for equations.
-
-\subsection{Phonetic fonts}
-
-We recommend that you use the \textsc{tipa} package for IPA phonetic
-symbols. For information about how to access Unicode fonts in the Word
-template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}. The font you use
-must be embedded. Please remember to check this, e.g.~by inspecting the
-``Document Properties --- Fonts'\,' in Acrobat Reader.
-
-\subsection{Page numbering}
-
-Page numbers will be added electronically to the document later.
-\textit{Please do not add page numbers and please do not insert
-any footers or headers!}
-
-\subsection{References}
-
-Please use just the reference number in square brackets. Formulations
-with author names like ``\ldots~as Ladefoged \cite{Ladefoged:2003}
-showed that \ldots'' are acceptable but not ``as shown in {[}Ladefoged,
-3{]}'' or ``as shown in (Ladefoged {[}3{]})''.
-
-References are to be numbered in alphabetical order. Please double-check
-the final version of your paper with regard to the correct
-correspondence of references to their numbers.
-
-\subsection{Hyperlinks}
-
-Links to URLs or email addresses should be formatted as normal text,
-\textit{not} as hyperlinks and not blue or underlined etc. Usually
-hyperlinks to web pages are listed in the references section. If
-required, line breaks can be placed within URLs after slashes or dashes
-(cf.\textasciitilde{}\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that
-no hyphens are inserted.
-
-\subsection{Footnotes and endnotes}
-
-If footnotes cannot be avoided they should appear as
-endnotes.\footnote{This footnote appears here as an endnote.}
-
-\section{Multimedia files}
-
-Multimedia data that are part of the paper are to be embedded in the
-submitted PDF; they cannot be submitted as supplementary data. Any
-images are to be included in the paper as Figures (see Section 2.3
-above). It is the authors' responsibility to check image quality ahead
-of submission. Audio examples are to be embedded within the PDF. To do
-this, authors can generate the PDF, and then embed the audio files using
-Adobe Acrobat Professional, or other software that offers the same
-outcome, so that the audio is included in the PDF. The presence of audio
-data should be identified in the text.
-
-We encourage authors to illustrate video data using still photographs
-from the video, and to include them as figures in the PDF. We cannot
-accept embedded video files, but authors are welcome to refer readers to
-a URL on the internet where these can be viewed.
-
-\section{PDF details}
-
-PDF files submitted must comply with the following requirements:
-
-\begin{enumerate}
-\item all special fonts and symbols must be embedded in the PDF
-  file so that correct rendering of the PDF does not depend on the
-  fonts installed on the viewer's computer
-\item there must be no
-  password protection on the PDF file, i.e. PDF files must not be
-  protected by PDF security in any way, i.e. content extraction,
-  document assembly, high resolution printing etc. must not be
-  forbidden
-\item PDF files should not contain any colours, hyperlinks,
-  multimedia or 3D content, and no JavaScript or forms
-\item PDF files
-  should be no larger than 5 MB.
-\end{enumerate}
-
-\section{Anonymity}
-
-In ICPhS 2023 submissions, an anonymous reviewing process will be used.
-This means that for the first submission the name(s) of the author(s)
-and their affiliation(s) \emph{must not} be mentioned. In addition,
-please refrain from using acknowledgements. Please also try to make your
-own previous research as anonymous as possible. As an example: do not
-write ``In our previous study \cite{Stevens:1999} we could show\ldots''
-but ``As shown in \cite{Stevens:1999}\ldots''. Or refer to your own
-published or otherwise widely known work, and to that of the other
-authors, in the ``Julius Caesar style'', i.e.~in the third person (for
-example: his work, her work, their work). Reference as ``anonymous''
-only work that you or the other authors have submitted for publication,
-but which has not yet been published, e.g.~\cite{Anonymous:submitted}.
-
-Please make sure that no author details appear in the Document
-Properties of the PDF file.
-\emph{For the revised paper submission author details are of course needed}.
-Acknowledgements and references to one's own work are possible as usual.
-
-\section{Format of references}
-
-References are formatted using the IEEE standard (available in various
-reference management systems like Zotero or Mendeley). If you do not use
-a reference management system, please use the examples provided for
-monographs \cite{Fant:1960}, contributions to volumes
-\cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles
-in conference proceedings \cite{Ladefoged:2003}. Abbreviations of
-well-known conferences and journals are possible
-\cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016}
-are referenced according to authors' instructions.
-
-\section{R scripts}
-
-You can use knitr code chunks like in any .Rmd document:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\DecValTok{2} \SpecialCharTok{+} \DecValTok{2}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 4
-\end{verbatim}
-
-This includes inline code: 2 + 2 = 4.
-
-R scripts can be loaded as well:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{source}\NormalTok{(}\StringTok{"./includes/scripts/analysis.R"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Which means you have access to any objects assigned in the script, like
-Figure \ref{fig:ex_plot}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cars\_plot}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[!h]
-\caption{This is a figure caption.}\label{fig:ex_plot}
-
-\begin{center}\includegraphics{main_files/figure-latex/ex_plot-1} \end{center}
-
-\end{figure}
-
 \bibliographystyle{./includes/bib/IEEEtran.bst}
@@
 
-\theendnotes
-
 \end{document}

### 2024-05-03 manuscript/main.Rmd

@@
 
-Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean coast **(Hualde, 1992)**.
-It is safe to assume that there is a bidirectional influence between these languages: in that region, Catalan has influenced Spanish and Spanish has influenced Catalan in all linguistic levels **(Davidson, 2014, 2015)**.
+Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean coast \cite{hualde_1992} It is safe to assume that there is a bidirectional influence between these languages: in that region, Catalan has influenced Spanish and Spanish has influenced Catalan in all linguistic levels \cite{davidson_2014} \cite{davidson_2015}.
 A crucial area of cross-linguistic influence might be the production of sounds, especially in those sounds that have different distribution in Catalan and Spanish.
-For example, previous studies have found that, in Barcelona, the production of Catalan vowels is different based on the L1 of the speaker, being it less target-like in L1 Spanish speakers compared to L1 Catalan speakers **(CITAR SEBASTIAN GALLES, etc)**.
+For example, previous studies have found that, in Barcelona, the production of Catalan vowels is different based on the L1 of the speaker, being it less target-like in L1 Spanish speakers compared to L1 Catalan speakers \cite{pallier_bosch_sebastian_galles_1997}.
 
 However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound.
-Therefore, the aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test **(Lambert, 1967)**.
-The alveolar fricative was chosen because it exists in both Catalan and Spanish inventories, but their contextual distribution is different **(Llisterri, 2023)**.
+Therefore, the aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test \cite{lambert_1967}.
+The alveolar fricative was chosen because it exists in both Catalan and Spanish inventories, but their contextual distribution is different \cite{llisterri_2023}.
 
@@
 
-Both Catalan and Peninsular Spanish sound inventories have a phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}.
-In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts **(Lloret, 2011; Recasens Vives, 1996, p.
-279)**.
-In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] **(Carrera-Sabaté & Prieto, 2012-2019)**.
-Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound **(Institut d’Estudis Catalans, 2022)**.
+Both Catalan and Peninsular Spanish sound inventories have a phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_2015}.
+In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts \cite{lloret_2011} \cite{recasens_1996}.
+In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] \cite{carrera_sabate_prieto_2012_2019}.
+Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound \cite{iec_2022}.
 For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].
 
-\cite{hualde_colina_2015
-
 While Spanish inventory does not have the phoneme /z/, the sound [z] may occur as an allophone of /s/, due to a voice assimilation process.
-It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.
-There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead.
-This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such as the so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
-268)**.
+It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial \cite{benet_cortes_lleo_2012} \cite{hualde_2015} \cite{hualde_et_al_2020}.
+There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules \cite{iec_2022}, the [z] should be produced, but [s] is produced instead.
+This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such as the so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas \cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020}.
 The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia.
-Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.
+Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia \cite{davidson_2020}.
 
@@
 
-The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual **(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003)** and bilingual contexts **(Davidson, 2014, 2015, 2019; Sinner, 2002)**, despite having been categorized as a low-salient variant **(Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)**.
-With respect to the former, it has been found that for listeners from Costa Rica and Ecuador (García, 2019), the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
+The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual \cite{chappel_et_al_2023} \cite{chappel_2021} \cite{garcia_2019} \cite{chappel_2016} \cite{chappell_2017} \cite{chappell_garcia_2017} \cite{silverstein_2003} and bilingual contexts \cite{davidson_2014} \cite{davidson_2015} \cite{davidson_2019} \cite{sinner_2002}, despite having been categorized as a low-salient variant \cite{chappel_et_al_2023} \cite{chappel_2021} \cite{chappell_2017} .
+With respect to the former, it has been found that for listeners from Costa Rica and Ecuador \cite{garcia_2019}, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender.
 Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine.
-Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class **(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003)**.
+Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class \cite{chappel_et_al_2023} \cite{chappel_2016} \cite{silverstein_2003}.
 Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects that exhibit systematic voicing.
 
-With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations **(Davidson, 2014, 2015)**.
-According to **Sinner (2002)**, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness **(Davidson, 2019; Sinner, 2002)**.
+With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations \cite{davidson_2014} \cite{davidson_2015}.
+According to Sinner \cite{sinner_2002}, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness \cite{davidson_2019} \cite{sinner_2002}.
 Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization.
 These results are in line with previous findings on intervocalic [z] in Catalonian Spanish.
-For instance, **Silverstein (2003)** found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride.
-Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness **(Davidson, 2019)**.
+For instance, Silverstein \cite{silverstein_2003} found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride.
+Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness \cite{davidson_2019}.
 
@@
 The first part, a demographic questionnaire, collected information regarding participant age, gender, place of origin, place of residence, mother tongue, family income, and level of studies.
-The second portion of the survey consisted of a perception task where participants were presented with a series of stimuli containing the target phoneme contrast (intervocalic voiced and voiceless alveolar fricative) such as *han tancat les vies estretes* 'they closed the narrow streets', followed by 11 evaluative statements related to social characteristics such as level of studies, friendliness, social class, and ethnic origin.
+The second portion of the survey consisted of a perception task where participants were presented with a series of stimuli containing the target phoneme contrast (intervocalic voiced and voiceless alveolar fricative) such as *Han tancat les vies estretes* 'they closed the narrow streets', followed by 11 evaluative statements related to social characteristics such as level of studies, friendliness, social class, and ethnic origin.
 These evaluative statements were then followed by a 100-point draggable sliding scale with which participants indicated their level of agreement or disagreement with the content of the statement, where 0 was “totally disagree” and 100 was “totally agree”.
@@
 
-The aim of this exploration study was to explore the covert attitudes that are linked to the perception Catalan intervocalic /s/ as either [s] or [z] and to examine the relationship between  individual social factors and covert attitudes towards Catalan intervocalic /s/.
+The aim of this exploratory study was to explore the covert attitudes that are linked to the perception Catalan intervocalic /s/ as either [s] or [z] and to examine the relationship between  individual social factors and covert attitudes towards Catalan intervocalic /s/.
 The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.
@@
 Contrary to what we expected, phoneme was not a significant predictor.
-One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] **(Davidson, 2019)** or, if they could, this difference in production was not changing their attitudes towards the speaker **(Sinner, 2002)**.
-This finding is interesting considering that the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p.
-268)**, and, from a prescriptivist point of view, it should be avoided **(Institut d'Estudis Catalans, 2022)**.
+One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] \cite{davidson_2019} or, if they could, this difference in production was not changing their attitudes towards the speaker \cite{sinner_2002}.
+This finding is interesting considering that the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish \cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020}, and, from a prescriptivist point of view, it should be avoided \cite{iec_2022}.
 Despite the lack of significance of this predictor, it should be mentioned that the speaker received more positive ratings when the [z] was being produced.
-The fact that phoneme was not an important factor contrasts with **Davidson (2020)**, who analyzed covert and overt attitudes towards Catalonian Spanish and found that [z] was a significant predictor: it was considered a sociolinguistic indicator of Catalan-Spanish bilingual membership.
-Although at this point we can only hypothesize, it could be that [z] is perceptually more salient in Spanish than in Catalan, as the Spanish inventory does not have the phoneme /z/, but voicing of /s/ might partially occur **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.
+The fact that phoneme was not an important factor contrasts with Davidson \cite{davidson_2020}, who analyzed covert and overt attitudes towards Catalonian Spanish and found that [z] was a significant predictor: it was considered a sociolinguistic indicator of Catalan-Spanish bilingual membership.
+Although at this point we can only hypothesize, it could be that [z] is perceptually more salient in Spanish than in Catalan, as the Spanish inventory does not have the phoneme /z/, but voicing of /s/ might partially occur \cite{benet_cortes_lleo_2012} \cite{hualde_2015} \cite{hualde_et_al_2020}.
 Another explanation could be that the /s/ alone does not provide enough social information; rather, the combination of the phoneme /s/ with the precedent vowel might give the listener more cues and have a stronger social meaning.
@@
 A potential explanation for this finding, that should be verified in future research, could be that L1 Catalan speakers have specific ideas (higher standards) towards the 'proper' Catalan.
-In addition, some studies have suggested that those Catalan people who grow up in Spanish-speaking households typically exhibit more negative attitudes towards Catalan **(CITATIONS)**, which is consistent with our results.
+In addition, some studies have suggested that those Catalan people who grow up in Spanish-speaking households typically exhibit more negative attitudes towards Catalan \cite{aguilo_more_lynch_2017}, which is consistent with our results.
 
@@
 We asked her to read the sentences using a Central Catalan 'neutral accent', which people might have potentially related with Barcelona.
-This 'neutral accent' does not show any of the prototypical characteristics of the *accent gironí*, a subvariety within Central Catalan spoken in some areas of Girona **(Gros, 2024)**.
+This 'neutral accent' does not show any of the prototypical characteristics of the *accent gironí*, a subvariety within Central Catalan spoken in some areas of Girona \cite{gros_2024}.
 

### 2024-05-03 manuscript/main.tex

@@
 Catalan and Spanish are two languages that have coexisted for centuries,
-mainly in eastern Spain, along the Mediterranean coast \textbf{(Hualde,
-1992)}. It is safe to assume that there is a bidirectional influence
-between these languages: in that region, Catalan has influenced Spanish
-and Spanish has influenced Catalan in all linguistic levels
-\textbf{(Davidson, 2014, 2015)}. A crucial area of cross-linguistic
-influence might be the production of sounds, especially in those sounds
-that have different distribution in Catalan and Spanish. For example,
-previous studies have found that, in Barcelona, the production of
-Catalan vowels is different based on the L1 of the speaker, being it
-less target-like in L1 Spanish speakers compared to L1 Catalan speakers
-\textbf{(CITAR SEBASTIAN GALLES, etc)}.
+mainly in eastern Spain, along the Mediterranean coast
+\cite{hualde_1992} It is safe to assume that there is a bidirectional
+influence between these languages: in that region, Catalan has
+influenced Spanish and Spanish has influenced Catalan in all linguistic
+levels \cite{davidson_2014} \cite{davidson_2015}. A crucial area of
+cross-linguistic influence might be the production of sounds, especially
+in those sounds that have different distribution in Catalan and Spanish.
+For example, previous studies have found that, in Barcelona, the
+production of Catalan vowels is different based on the L1 of the
+speaker, being it less target-like in L1 Spanish speakers compared to L1
+Catalan speakers \cite{pallier_bosch_sebastian_galles_1997}.
 
@@
 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched
-guise test \textbf{(Lambert, 1967)}. The alveolar fricative was chosen
+guise test \cite{lambert_1967}. The alveolar fricative was chosen
 because it exists in both Catalan and Spanish inventories, but their
-contextual distribution is different \textbf{(Llisterri, 2023)}.
+contextual distribution is different \cite{llisterri_2023}.
 
@@
 /s/, which can be classified as a voiceless alveolar fricative
-\cite{hualde_colina_2015}. In addition, most of the Catalan varieties
-have a voiced alveolar phoneme in intervocalic contexts \textbf{(Lloret,
-2011; Recasens Vives, 1996, p.~279)}. In onset position, /s/ and /z/
-contrast phonemically, which leads to minimal pairs like \emph{cel}
+\cite{hualde_2015}. In addition, most of the Catalan varieties have a
+voiced alveolar phoneme in intervocalic contexts \cite{lloret_2011}
+\cite{recasens_1996}. In onset position, /s/ and /z/ contrast
+phonemically, which leads to minimal pairs like \emph{cel}
 {[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
-\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} \textbf{(Carrera-Sabaté
-\& Prieto, 2012-2019)}. Conversely, at the end of the word, this
-contrast is neutralized, being the voicing or devoicing of the alveolar
-fricative dependent on the following sound \textbf{(Institut d'Estudis
-Catalans, 2022)}. For example, if it is followed by a vowel, the
-alveolar fricative is voiced: \emph{torres} {[}tór\textschwa{}s{]} and
-\emph{torres antigues}
+\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]}
+\cite{carrera_sabate_prieto_2012_2019}. Conversely, at the end of the
+word, this contrast is neutralized, being the voicing or devoicing of
+the alveolar fricative dependent on the following sound \cite{iec_2022}.
+For example, if it is followed by a vowel, the alveolar fricative is
+voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
 {[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.
-
-\textbackslash cite\{hualde\_colina\_2015
 
@@
 It is important to mention, though, that it has been claimed that this
-voicing is generally not complete, but partial \textbf{(Benet et al.,
-2012; Hualde \& Colina, 2015; Hualde et al., 2020)}. There is some
-evidence of the influence of Spanish on the production of the Catalan
-alveolar fricative, in contexts where, according to prescriptivist rules
-\textbf{(Institut d'Estudis Catalans, 2022)}, the {[}z{]} should be
-produced, but {[}s{]} is produced instead. This phenomenon has mainly
-been attested in those areas where there has been an extended contact
-between Catalan and Spanish, such as the so-called ``franja de Aragón''
-(borderland between Catalonia and Aragon), some areas of the Valencian
-Community, and Barcelona and other urban areas \textbf{(Benet et al.,
-2012; Davidson, 2020; Recasens Vives, 1996, p.~268)}. The opposite
-pattern has also been observed - that is, the voicing of the alveolar
-fricative in the Spanish spoken in Barcelona and Valencia.
+voicing is generally not complete, but partial
+\cite{benet_cortes_lleo_2012} \cite{hualde_2015}
+\cite{hualde_et_al_2020}. There is some evidence of the influence of
+Spanish on the production of the Catalan alveolar fricative, in contexts
+where, according to prescriptivist rules \cite{iec_2022}, the {[}z{]}
+should be produced, but {[}s{]} is produced instead. This phenomenon has
+mainly been attested in those areas where there has been an extended
+contact between Catalan and Spanish, such as the so-called ``franja de
+Aragón'' (borderland between Catalonia and Aragon), some areas of the
+Valencian Community, and Barcelona and other urban areas
+\cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020}.
+The opposite pattern has also been observed - that is, the voicing of
+the alveolar fricative in the Spanish spoken in Barcelona and Valencia.
 Interestingly, although in both Barcelona and Valencia Catalan and
@@
 greater influence on Spanish in Barcelona, but Spanish has a greater
-influence on Catalan in Valencia \textbf{(Davidson, 2020)}.
+influence on Catalan in Valencia \cite{davidson_2020}.
 
@@
 has been studied in a variety of contexts, including both monolingual
-\textbf{(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \&
-García, 2017; García, 2015, 2019; Silverstein, 2003)} and bilingual
-contexts \textbf{(Davidson, 2014, 2015, 2019; Sinner, 2002)}, despite
-having been categorized as a low-salient variant \textbf{(Chappell,
-2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
-former, it has been found that for listeners from Costa Rica and Ecuador
-(García, 2019), the covert attitudes associated with intervocalic
-{[}z{]} vary significantly by speaker gender. Consequently, male
-speakers who use {[}z{]} tend to be positively evaluated, and generally
-perceived as nice, local, confident, and masculine. Conversely, female
-speakers who use {[}z{]} are only afforded negative associations of
-lower education and social class \textbf{(Chappell, 2016; Chappell et
-al., 2023; Silverstein, 2003)}. Additionally, this variable has been
-found to serve as a marker of regional identity, with voicing being
-associated with neighboring Highland Ecuadorian dialects that exhibit
-systematic voicing.
+\cite{chappel_et_al_2023} \cite{chappel_2021} \cite{garcia_2019}
+\cite{chappel_2016} \cite{chappell_2017} \cite{chappell_garcia_2017}
+\cite{silverstein_2003} and bilingual contexts \cite{davidson_2014}
+\cite{davidson_2015} \cite{davidson_2019} \cite{sinner_2002}, despite
+having been categorized as a low-salient variant
+\cite{chappel_et_al_2023} \cite{chappel_2021} \cite{chappell_2017} .
+With respect to the former, it has been found that for listeners from
+Costa Rica and Ecuador \cite{garcia_2019}, the covert attitudes
+associated with intervocalic {[}z{]} vary significantly by speaker
+gender. Consequently, male speakers who use {[}z{]} tend to be
+positively evaluated, and generally perceived as nice, local, confident,
+and masculine. Conversely, female speakers who use {[}z{]} are only
+afforded negative associations of lower education and social class
+\cite{chappel_et_al_2023} \cite{chappel_2016} \cite{silverstein_2003}.
+Additionally, this variable has been found to serve as a marker of
+regional identity, with voicing being associated with neighboring
+Highland Ecuadorian dialects that exhibit systematic voicing.
 
@@
 {[}z{]} in bilingual contexts, this phenomenon has been mainly studied
-in Catalan-Spanish contact situations \textbf{(Davidson, 2014, 2015)}.
-According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
-considered a possible linguistic marker of Catalonian Spanish, despite
-its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
-2002)}. Additionally, Davidson (2019) found that, in Barcelona, the
-indexical field of intervocalic {[}z{]} features exclusively positive
-associations with bilingualism and local solidarity, which warrants its
-gradual adoption in the speech community, unlike other phonetic markers
-of Catalonian Spanish that are the topic of overt social ridicule, such
-as lateral velarization. These results are in line with previous
-findings on intervocalic {[}z{]} in Catalonian Spanish. For instance,
-\textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
-local Barcelona identity, and the variant then acquires positive
-associations with bilingualism and cultural pride. Overall, these
-findings suggest that intervocalic {[}z{]} likely functions as a
-sociolinguistic indicator of Catalonian Spanish within the Catalonian
-speech community, indexing Catalan-Spanish bilingual membership while
-lacking any considerable degree of generalized conscious awareness
-\textbf{(Davidson, 2019)}.
+in Catalan-Spanish contact situations \cite{davidson_2014}
+\cite{davidson_2015}. According to Sinner \cite{sinner_2002},
+intervocalic {[}z{]} can be considered a possible linguistic marker of
+Catalonian Spanish, despite its relative lack of social awareness
+\cite{davidson_2019} \cite{sinner_2002}. Additionally, Davidson (2019)
+found that, in Barcelona, the indexical field of intervocalic {[}z{]}
+features exclusively positive associations with bilingualism and local
+solidarity, which warrants its gradual adoption in the speech community,
+unlike other phonetic markers of Catalonian Spanish that are the topic
+of overt social ridicule, such as lateral velarization. These results
+are in line with previous findings on intervocalic {[}z{]} in Catalonian
+Spanish. For instance, Silverstein \cite{silverstein_2003} found that
+intervocalic {[}z{]} indexes a local Barcelona identity, and the variant
+then acquires positive associations with bilingualism and cultural
+pride. Overall, these findings suggest that intervocalic {[}z{]} likely
+functions as a sociolinguistic indicator of Catalonian Spanish within
+the Catalonian speech community, indexing Catalan-Spanish bilingual
+membership while lacking any considerable degree of generalized
+conscious awareness \cite{davidson_2019}.
 
@@
 the target phoneme contrast (intervocalic voiced and voiceless alveolar
-fricative) such as \emph{han tancat les vies estretes} `they closed the
+fricative) such as \emph{Han tancat les vies estretes} `they closed the
 narrow streets', followed by 11 evaluative statements related to social
@@
 
-The aim of this exploration study was to explore the covert attitudes
+The aim of this exploratory study was to explore the covert attitudes
 that are linked to the perception Catalan intervocalic /s/ as either
@@
 participants could not distinguish between {[}s{]} and {[}z{]}
-\textbf{(Davidson, 2019)} or, if they could, this difference in
-production was not changing their attitudes towards the speaker
-\textbf{(Sinner, 2002)}. This finding is interesting considering that
-the production of the voiceless alveolar fricative in intervocalic
-position has been considered a result of Catalan in contact with Spanish
-\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
-p.~268)}, and, from a prescriptivist point of view, it should be avoided
-\textbf{(Institut d'Estudis Catalans, 2022)}. Despite the lack of
-significance of this predictor, it should be mentioned that the speaker
-received more positive ratings when the {[}z{]} was being produced. The
-fact that phoneme was not an important factor contrasts with
-\textbf{Davidson (2020)}, who analyzed covert and overt attitudes
-towards Catalonian Spanish and found that {[}z{]} was a significant
-predictor: it was considered a sociolinguistic indicator of
-Catalan-Spanish bilingual membership. Although at this point we can only
-hypothesize, it could be that {[}z{]} is perceptually more salient in
-Spanish than in Catalan, as the Spanish inventory does not have the
-phoneme /z/, but voicing of /s/ might partially occur \textbf{(Benet et
-al., 2012; Hualde \& Colina, 2015; Hualde et al., 2020)}. Another
-explanation could be that the /s/ alone does not provide enough social
-information; rather, the combination of the phoneme /s/ with the
-precedent vowel might give the listener more cues and have a stronger
-social meaning. Future studies should address this issue in order to
-explore if the combination of vowel + /s/ carries other social
-connotations than just the /s/ alone.
+\cite{davidson_2019} or, if they could, this difference in production
+was not changing their attitudes towards the speaker \cite{sinner_2002}.
+This finding is interesting considering that the production of the
+voiceless alveolar fricative in intervocalic position has been
+considered a result of Catalan in contact with Spanish
+\cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020},
+and, from a prescriptivist point of view, it should be avoided
+\cite{iec_2022}. Despite the lack of significance of this predictor, it
+should be mentioned that the speaker received more positive ratings when
+the {[}z{]} was being produced. The fact that phoneme was not an
+important factor contrasts with Davidson \cite{davidson_2020}, who
+analyzed covert and overt attitudes towards Catalonian Spanish and found
+that {[}z{]} was a significant predictor: it was considered a
+sociolinguistic indicator of Catalan-Spanish bilingual membership.
+Although at this point we can only hypothesize, it could be that {[}z{]}
+is perceptually more salient in Spanish than in Catalan, as the Spanish
+inventory does not have the phoneme /z/, but voicing of /s/ might
+partially occur \cite{benet_cortes_lleo_2012} \cite{hualde_2015}
+\cite{hualde_et_al_2020}. Another explanation could be that the /s/
+alone does not provide enough social information; rather, the
+combination of the phoneme /s/ with the precedent vowel might give the
+listener more cues and have a stronger social meaning. Future studies
+should address this issue in order to explore if the combination of
+vowel + /s/ carries other social connotations than just the /s/ alone.
 
@@
 Spanish-speaking households typically exhibit more negative attitudes
-towards Catalan \textbf{(CITATIONS)}, which is consistent with our
-results.
+towards Catalan \cite{aguilo_more_lynch_2017}, which is consistent with
+our results.
 
@@
 \emph{accent gironí}, a subvariety within Central Catalan spoken in some
-areas of Girona \textbf{(Gros, 2024)}.
+areas of Girona \cite{gros_2024}.
 
